### 2019-05-09 manuscript/outline.Rmd
outline

---
output: 
  word_document:
    reference_docx: "common/word-styles-reference-01.docx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Combining genomic prediction with citizen science for crop improvement, as case study with “early selection” in durum wheat

# Authors:

Kauê de Sousa, Matteo Dell’Acqua, Carlo Fadda, Yosef Gebrehawaryat, Jean-Luc Jannink, Dejene Mengistu, Mario Enrico Pè, [… add others], Jacob van Etten, Svein Ø. Solberg

# What we have:

1.	Station data for 400 farmer varieties (genebank accession) and a few modern varieties (derived from Mediterranean germplasm) in two regions of Ethiopia
2.	Citizen science (tricot) data for > 60 varieties (a subset from the 400) in 3 seasons with >1k plots in total, 3 varieties + check for each plot (Calculate how many times each variety is repeated)
    i) Each observation has a rank of 4 randomised varieties, with planting dates, geographic location and some other information, such number of seeds per spike and reason for the rank
3. High-density SNP genomic data for 400 varieties, using Illumina

# Research questions:

- Is genomic selection with citizen science a viable approach?

- Does the genetic information provide sufficient gain in the predictions for variety selection (given the cost)?

- Can the combination of genomic and citizen science provide new insights in which traits farmers prioritise or which areas of the genome play a role?

- Does genomic prediction using phenotypic data produced on station (agronomic traits, farmer preferences) improve accuracy in the field as measured in the tricot trial?

# Analytical/Model approaches

Run a number of scenarios from varieties selection using the citizen science and station data to determine whether the genetic information provides any gain or not

## Scenario 1

Baseline: tricot without genomic prediction, blocked cross-validation with seasons as blocks

- Plackett-Luce model

- Bradley-Terry model. Improvement: same exercise, but adding genomic data as explanatory data

- Plackett-Luce model with covariance matrix derived from genomic data

- Bradley-Terry model with genomic random variables in a GLM framework(?)

## Scenario 2

Baseline: farmer preferences from station rating exercise, predicting field performance in tricot perhaps stratifying by environment and using the two station sites separately for their respective most similar environments (to circumvent GxE).

Predict rating data with an appropriate GLM model for ordinal data (the rates should not be assumed to be equidistant, the difference between rates 1 and 2 may be different from the rates between 4 and 5).

Improvement: same farmer preferences + genomic data, letting the ordering be influenced by genomic relatedness
The exercise could use rank correlation, Kendall´s tau as criterion, but could also make a real “selection” (say top 50%) of the accessions and determine the gain as the % that indeed falls in the top 50% in the tricot trial. A p value for this exercise could be determined by doing 10,000 random selections of 50% to get a null distribution.

# Topics to explore:

- What is the relative costs of these approaches

- Is data missingness non-random? Can data missingness be predicted from farmer preferences?

- Can we say something about which genomic regions play a role?

### 2019-05-27 manuscript/manuscript.Rmd
25052018

---
title: "Combining genomic prediction with citizen science for crop improvement and early selection in durum wheat"
bibliography: "library/tricotgenomic.bib"
csl: "library/citation-style.csl"
output: 
  word_document:
    reference_docx: "library/word-styles-reference-01.docx"
# output: pdf_document
---

Kauê de Sousa^1,2,^*, Matteo Dell’Acqua^3^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Dejene Mengistu^7^, Mario Enrico Pè^3^, [… add others], Jacob van Etten^2^, Svein Ø. Solberg^1^  
^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
^3^ Scuola Superiore Sant’Anna, Institute of Life Sciences, 56124 Pisa, Italy  
^4^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
^5^ College of Agriculture and Life Sciences, Cornell University, 14853 Ithaca, USA  
^6^ Agricultural Research Service, United States Department of Agriculture, 148532901, Ithaca, USA  
^7^ Bioversity International, Nepal Office, 3055 Kathmandu, Nepal

**Research questions:**

* Is genomic selection with citizen science a viable approach?
* Does the genetic information provide sufficient gain in the predictions for variety selection (given the cost)?
* Can the combination of genomic and citizen science provide new insights in which traits farmers prioritise or which areas of the genome play a role?
* Does genomic prediction using phenotypic data produced on station (agronomic traits, farmer preferences) improve accuracy in the field as measured in the tricot trial?

# Introduction

* Crowdsourcing citizen science showed the ability to register the climate effects as they occurr in farms, with potential to be used in plant breeding programs and risk assessments for farm adaptation[@VanEtten2019].

# Methods

## Model approaches

Run a number of scenarios from varieties selection using the citizen science and station data to determine whether the genetic information provides any gain or not

**Scenario 1**

Baseline: tricot without genomic prediction, blocked cross-validation with seasons as blocks

* Plackett-Luce model
* Bradley-Terry model. Improvement: same exercise, but adding genomic data as explanatory data
* Plackett-Luce model with covariance matrix derived from genomic data
* Bradley-Terry model with genomic random variables in a GLM framework(?)

**Scenario 2**

Baseline: farmer preferences from station rating exercise, predicting field performance in tricot perhaps stratifying by environment and using the two station sites separately for their respective most similar environments (to circumvent GxE).

Predict rating data with an appropriate GLM model for ordinal data (the rates should not be assumed to be equidistant, the difference between rates 1 and 2 may be different from the rates between 4 and 5).

Improvement: same farmer preferences + genomic data, letting the ordering be influenced by genomic relatedness
The exercise could use rank correlation, Kendall´s tau as criterion, but could also make a real “selection” (say top 50%) of the accessions and determine the gain as the % that indeed falls in the top 50% in the tricot trial. A p value for this exercise could be determined by doing 10,000 random selections of 50% to get a null distribution.

# Results

# Discussion

* What is the relative costs of these approaches
* Is data missingness non-random? Can data missingness be predicted from farmer preferences? 
* Can we say something about which genomic regions play a role?

# References

### 2019-05-27 manuscript/outline.Rmd
add_data
25052018

@@
 ---
+title: "Combining genomic prediction with citizen science for crop improvement, as case study with “early selection” in durum wheat"
 output: 
   word_document:
-    reference_docx: "common/word-styles-reference-01.docx"
+    reference_docx: "library/word-styles-reference-01.docx"
 ---
@@
 ```
-
-# Combining genomic prediction with citizen science for crop improvement, as case study with “early selection” in durum wheat
 

### 2019-05-31 script/01_prepare_data.R
data_cleaning

@@ -1,0 +1,329 @@
+####################################################
+###### Read and clean durum wheat in Ethiopia
+# Updated 31May2019
+####################################################
+
+library("tidyverse")
+library("readxl")
+library("magrittr")
+library("janitor")
+library("sf")
+
+# ..........................................
+# ..........................................
+# file with some pre-debuged variables ####
+df <- "data/raw/climmob_ethiopia_2017.csv"
+df %<>% 
+  read_csv(., 
+           na = c("#VALUE!", "NA", "<NA>", ""))
+
+# tidy up the colnames
+df %<>% 
+  as_tibble(.name_repair = janitor::make_clean_names) %>%
+  rename(lon = longtiude,
+         lat = latitude)
+
+# drop all missing rankings or rankings with missing Accession (item id)
+df %<>%
+  filter(!is.na(farmer_rank) & !is.na(accession))
+
+# only keep strict rankings of at least 2 distinct items
+df %>%
+  group_by(farmer_no) %>%
+  summarise(keep = setequal(farmer_rank, 1:4) & n_distinct(accession) == 4) %>%
+  filter(keep) ->
+  keep
+
+# apply the logical vector
+id <- df$farmer_no %in% keep$farmer_no
+
+# take rankings
+n <- nrow(keep)
+
+itemnames <- sort(unique(df$accession[id]))
+
+# keep selected observations
+df <- df[id,]
+
+# drop some variables
+names(df)
+drop <- c("no" , "code_region" , "code_year" , "code_zone" , "code_district",
+          "code_kebelle" , "altitude" , "code_farmer" , "sex" , "code_accession",
+          "days2maturity" , "gy_gm" , "biomass_gm")
+
+drop <- !names(df) %in% drop
+
+df <- df[drop]
+
+
+# rename variety Hetosa
+df %<>% 
+  mutate(accession = ifelse(accession == "Hetosa",
+                            "Hitosa", 
+                            accession))
+
+# reshape values
+df %>% 
+  select(farmer_no, plot_no, accession) %>% 
+  spread(.,
+         key = plot_no,
+         value = "accession") ->
+  vars 
+  
+names(vars)[2:5] <- paste0("variety_", letters[1:4])
+
+df %>% 
+  select(farmer_no, plot_no,farmer_rank) %>% 
+  spread(.,
+         key = plot_no,
[…]

### 2019-05-31 script/03_model.R

@@ -1,0 +1,25 @@
+library(gosset)
+library(PlackettLuce)
+library(BradleyTerryScalable)
+library(igraph)
+
+R <- to_rankings(items = df[paste0("variety_",letters[1:4])],
+                 input = df[paste0("rank_variety_",letters[1:4])])
+
+
+adj <- PlackettLuce::adjacency(R)
+
+adj <- as.vector(adj)
+
+adj <- t(matrix(adj, nrow = ncol(R), ncol = ncol(R)))
+
+dimnames(adj) <- list(dimnames(R)[[2]], dimnames(R)[[2]])
+
+adj <- btdata(adj, return_graph = TRUE)
+
+plot.igraph(adj$graph, vertex.size = 10, edge.arrow.size = 0.1) 
+
+
+mod <- PlackettLuce(R)
+summary(mod)
+plot(qvcalc(mod), las = 2, cex.axis = 0.7)

### 2019-05-31 manuscript/manuscript.Rmd
passport_data

@@
 
-Kauê de Sousa^1,2,^*, Matteo Dell’Acqua^3^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Dejene Mengistu^7^, Mario Enrico Pè^3^, [… add others], Jacob van Etten^2^, Svein Ø.
+Kauê de Sousa^1,2,^*, Matteo Dell’Acqua^3^, Jacob van Etten^2^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Dejene Mengistu^7^, Mario Enrico Pè^3^, [… add others], Svein Ø.
 Solberg^1^

### 2019-06-09 script/01_prepare_data.R
input

@@ -1,28 +1,32 @@
 ####################################################
 ###### Read and clean durum wheat in Ethiopia
-# Updated 06Jun2019
+# Updated 07Jun2019
 ####################################################
 
 library("tidyverse")
-library("readxl")
 library("magrittr")
+library("readxl")
 library("janitor")
 library("sf")
 
 # ..........................................
 # ..........................................
 # file with some pre-debuged variables ####
-df <- "data/raw/climmob_ethiopia_2017.csv"
+df <- "data/raw/DataSheet_CS_All_FINAL.csv"
 df %<>% 
-  read_csv(., 
-           na = c("#VALUE!", "NA", "<NA>", ""))
+  read.csv(.,
+           na.strings = c("NA", " ", "missing","?","#DIV/0!","#REF!")) %>% 
+  as_tibble()
+
+
+df
 
 # tidy up the colnames
 df %<>% 
   as_tibble(.name_repair = janitor::make_clean_names) %>%
   rename(lon = longtiude,
-         lat = latitude) %>% 
-  select(-genotype)
+         lat = latitude)
+
 
 # drop all missing rankings or rankings with missing Accession (item id)
 df %<>%
@@ -30,15 +34,15 @@ df %<>%
 
 # drop some variables
 names(df)
-drop <- c("no" , "code_region" , "code_year" , "code_zone" , "code_district",
-          "code_kebelle" , "altitude" , "code_farmer" , "sex" , "code_accession",
-          "days2maturity" , "gy_gm" , "biomass_gm")
+drop <- c("no" , "altitude" , "sex" , "code_accession",
+          "days2maturity" )
+
 
 drop <- !names(df) %in% drop
 
 df <- df[drop]
 
-# rename variety Hetosa
+# rename variety Hetosa and Assasa
 df %<>% 
   mutate(accession = ifelse(accession == "Hetosa",
                             "Hitosa", 
@@ -50,110 +54,6 @@ df %<>%
 
 # ..........................................
 # ..........................................
-# Drop variables not genotyped ####
-out <- "data/raw/not_in_genotyping.csv"
-out %<>% 
-  read_csv() %>% 
-  select(accession) %>% 
-  t()
-
-
-out <- !df$accession %in% out
-
-# retain the observations with genotype data
-df %<>% 
-  filter(out)
-
-# add genotype codes
[…]

### 2019-06-09 script/02_add_climate_indices.R

@@ -14,6 +14,11 @@ df <- "data/durumwheat.csv"
 df %<>% 
   read_csv()
 
+# keep only unique values for farmer id 
+df %<>% 
+  distinct(id, .keep_all = TRUE) %>% 
+  dplyr::select(id, planting_date, lon, lat)
+
 n <- nrow(df)
 
 # define some parameters
@@ -36,7 +41,7 @@ dates <- NULL
 summary(df$planting_date)
 
 # generate a sequency of 8-days from 2013 to 2015
-for(i in c(2003:2015)){
+for(i in c(2012:2015)){
   dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), 
                           as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
 }
@@ -60,11 +65,6 @@ modis <- array(NA, dim = c(n, length(dates), 2),
 # files in this directory has different extensions, 
 # so each tile is loaded individually
 # they are also very large so it is retrieved from a external drive
-# check if the drive exists
-
-if (!exists(edrive)) {
-  stop("An external drive with MODIS data is not available")
-}
 
 for (i in 1:2){
   
@@ -110,40 +110,45 @@ for (i in 1:dim(modis)[2]) {
 
 # apply linear interpolation to the 8-day temperature
 modis_approx <- array(NA, c(n, length(full_dates), 2), 
-                      dimnames = list(mydata$id,  as.character(full_dates), c(1:2) ))
+                      dimnames = list(df$id,  
+                                      as.character(full_dates), c(1:2) ))
 
 for(i in 1:dim(modis)[1]){
   modis_approx[i,,1] <- approx(modis[i,,1], n = length(full_dates))[[2]]
   modis_approx[i,,2] <- approx(modis[i,,2], n = length(full_dates))[[2]]
 }
-#save(modis, modis_approx, file = "./processing/modis_ethiopia.RData")
-load("./processing/modis_ethiopia.RData")
+
+save(modis, modis_approx, file = "./processing/modis.rda")
+load("./processing/modis.rda")
 
 
 # #add CHIRPS data
 cat("Readind CHIRPS data. Time:", date(), "\n")
-# #source CHIRPS (precipitation data)
-# #http://chg.geog.ucsb.edu/data/chirps/
-# years <- years <- seq(from=2003 , to=2017, by=1)
-# chirps <- list()
[…]

### 2019-06-09 script/03_model.R

@@ -112,6 +112,7 @@ kendall <- kendallTau(FR, YR)
 
 write_csv(kendall, paste0(output, "kendall_correlation.csv"))
 
+
 # dimnames(R)[2]
 # 
 # adj <- PlackettLuce::adjacency(R)

### 2019-06-09 script/01_prepare_data.R
data

@@ -1,11 +1,10 @@
 ####################################################
 ###### Read and clean durum wheat in Ethiopia
-# Updated 07Jun2019
+# Updated 09un2019
 ####################################################
 
 library("tidyverse")
 library("magrittr")
-library("readxl")
 library("janitor")
 library("sf")
 
@@ -15,28 +14,21 @@ library("sf")
 df <- "data/raw/DataSheet_CS_All_FINAL.csv"
 df %<>% 
   read.csv(.,
-           na.strings = c("NA", " ", "missing","?","#DIV/0!","#REF!")) %>% 
-  as_tibble()
-
-
-df
-
-# tidy up the colnames
-df %<>% 
+           na.strings = c("NA","", " ", "missing","?","#DIV/0!","#REF!")) %>% 
   as_tibble(.name_repair = janitor::make_clean_names) %>%
   rename(lon = longtiude,
-         lat = latitude)
+         lat = latitude) %>% 
+  mutate_if(is.factor, as.character)
 
 
-# drop all missing rankings or rankings with missing Accession (item id)
+# drop all missing rankings or rankings with missing accession
 df %<>%
   filter(!is.na(farmer_rank) & !is.na(accession))
 
 # drop some variables
 names(df)
-drop <- c("no" , "altitude" , "sex" , "code_accession",
-          "days2maturity" )
 
+drop <- c("no","altitude","sex","code_accession","days2maturity")
 
 drop <- !names(df) %in% drop
 
@@ -50,7 +42,7 @@ df %<>%
          accession = ifelse(accession == "Assassa",
                             "Asassa", 
                             accession))
-
+summary(as.factor(df$accession))
 
 # ..........................................
 # ..........................................
@@ -66,19 +58,19 @@ df %<>%
 raster::getData("GADM", country = "ETH", path = "data", level = 0) %>% 
   st_as_sf() -> eth
 
-plot(eth["GID_0"], col = "lightgrey")
-
-df %>% 
-  select(lon, lat) %>%  
-  filter(!is.na(lon)) %>% 
-  as.matrix() %>% 
-  st_multipoint() %>%
-  st_sfc(crs = st_crs(eth)) -> 
-  lonlat
-
-plot(eth["GID_0"], col = "lightgrey", reset = FALSE)
-plot(lonlat, col = "blue", cex = 1, 
[…]

### 2019-06-09 script/02_add_climate_indices.R

@@ -14,11 +14,6 @@ df <- "data/durumwheat.csv"
 df %<>% 
   read_csv()
 
-# keep only unique values for farmer id 
-df %<>% 
-  distinct(id, .keep_all = TRUE) %>% 
-  dplyr::select(id, planting_date, lon, lat)
-
 n <- nrow(df)
 
 # define some parameters
@@ -41,7 +36,7 @@ dates <- NULL
 summary(df$planting_date)
 
 # generate a sequency of 8-days from 2013 to 2015
-for(i in c(2012:2015)){
+for(i in c(2003:2015)){
   dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), 
                           as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
 }
@@ -65,6 +60,11 @@ modis <- array(NA, dim = c(n, length(dates), 2),
 # files in this directory has different extensions, 
 # so each tile is loaded individually
 # they are also very large so it is retrieved from a external drive
+# check if the drive exists
+
+if (!exists(edrive)) {
+  stop("An external drive with MODIS data is not available")
+}
 
 for (i in 1:2){
   
@@ -110,45 +110,40 @@ for (i in 1:dim(modis)[2]) {
 
 # apply linear interpolation to the 8-day temperature
 modis_approx <- array(NA, c(n, length(full_dates), 2), 
-                      dimnames = list(df$id,  
-                                      as.character(full_dates), c(1:2) ))
+                      dimnames = list(mydata$id,  as.character(full_dates), c(1:2) ))
 
 for(i in 1:dim(modis)[1]){
   modis_approx[i,,1] <- approx(modis[i,,1], n = length(full_dates))[[2]]
   modis_approx[i,,2] <- approx(modis[i,,2], n = length(full_dates))[[2]]
 }
-
-save(modis, modis_approx, file = "./processing/modis.rda")
-load("./processing/modis.rda")
+#save(modis, modis_approx, file = "./processing/modis_ethiopia.RData")
+load("./processing/modis_ethiopia.RData")
 
 
 # #add CHIRPS data
 cat("Readind CHIRPS data. Time:", date(), "\n")
-#source of CHIRPS (precipitation data)
-#http://chg.geog.ucsb.edu/data/chirps/
-years <- years <- seq(from = 2012 , to = 2016, by = 1)
-chirps <- list()
[…]

### 2019-06-09 script/03_model.R

@@ -38,6 +38,8 @@ G <- cbind(G, expvar)
 
 mod <- pltree(G ~ ., data = G)
 
+plot(mod)
+
 winprobs <- itempar(mod, log = TRUE)
 
 winprobs <- bind_cols(genotype = dimnames(winprobs)[[2]],
@@ -111,7 +113,6 @@ FR <- FR[1:length(FR), , as.grouped_rankings = FALSE]
 kendall <- kendallTau(FR, YR)
 
 write_csv(kendall, paste0(output, "kendall_correlation.csv"))
-
 
 # dimnames(R)[2]
 # 

### 2019-06-10 script/02_add_climate_indices.R
modis_chirps

@@ -14,6 +14,9 @@ df <- "data/durumwheat.csv"
 df %<>% 
   read_csv()
 
+df %<>%
+  distinct(id, .keep_all = TRUE)
+
 n <- nrow(df)
 
 # define some parameters
@@ -31,12 +34,12 @@ k_to_c <- 273.15
 
 tiles <- c("h21v07","h21v08")
 
+years <- c(2012:2016)
+
 dates <- NULL
 
-summary(df$planting_date)
-
 # generate a sequency of 8-days from 2013 to 2015
-for(i in c(2003:2015)){
+for(i in c(2012:2015)){
   dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), 
                           as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
 }
@@ -62,10 +65,6 @@ modis <- array(NA, dim = c(n, length(dates), 2),
 # they are also very large so it is retrieved from a external drive
 # check if the drive exists
 
-if (!exists(edrive)) {
-  stop("An external drive with MODIS data is not available")
-}
-
 for (i in 1:2){
   
   if(i == 1 ) m <- "day" else m <- "night"
@@ -75,6 +74,10 @@ for (i in 1:2){
     file <- list.files(paste0(edrive, "modis/smooth_Ethiopia/", tiles[j]), 
                        pattern = m, 
                        full.names = TRUE)
+    
+    
+    file <- file[grepl(paste0(years, collapse = "|"), file)]
+    
     
     file <- stack(file)
     
@@ -110,45 +113,38 @@ for (i in 1:dim(modis)[2]) {
 
 # apply linear interpolation to the 8-day temperature
 modis_approx <- array(NA, c(n, length(full_dates), 2), 
-                      dimnames = list(mydata$id,  as.character(full_dates), c(1:2) ))
+                      dimnames = list(df$id,  as.character(full_dates), c(1:2) ))
 
 for(i in 1:dim(modis)[1]){
   modis_approx[i,,1] <- approx(modis[i,,1], n = length(full_dates))[[2]]
   modis_approx[i,,2] <- approx(modis[i,,2], n = length(full_dates))[[2]]
 }
-#save(modis, modis_approx, file = "./processing/modis_ethiopia.RData")
-load("./processing/modis_ethiopia.RData")
-
+#save(modis, modis_approx, file = "data/modis.rda")
 
[…]

### 2019-06-10 script/helper_02_fetch_modis_chirps.R

@@ -1,0 +1,152 @@
+# Fetch MODIS and CHIRPS data using gps points 
+# 
+library("tidyverse")
+library("magrittr")
+library("raster")
+
+
+#.......................................
+#.......................................
+# Read data ####
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv()
+
+df %<>%
+  distinct(id, .keep_all = TRUE)
+
+n <- nrow(df)
+
+# define some parameters
+
+# the external drive with MODIS and CHIRPS data
+edrive <- "E:/rasters/"
+
+my_proj <- "+proj=longlat +datum=WGS84"
+
+ext <- raster::extent(68, 86, 9, 33)
+
+
+#.......................................
+#.......................................
+# Get MODIS ####
+
+modis_factor <- 0.02
+
+k_to_c <- 273.15
+
+tiles <- c("h21v07","h21v08")
+
+years <- c(2012:2016)
+
+dates <- NULL
+
+# generate a sequency of 8-days from 2013 to 2015
+for(i in c(2012:2015)){
+  dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), 
+                          as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
+}
+
+dates <- as.Date(c(dates, seq(as.Date("2016-01-01", "%Y-%m-%d"), 
+                              as.Date("2016-12-26", "%Y-%m-%d"), 8)), 
+                 origin = "1970-01-01")
+
+# and a vector with full dates (1 days)
+full_dates <- seq(dates[1], dates[length(dates)], 1)
+
+# create an array with 2 layers
+# modis day and night
+modis <- array(NA, dim = c(n, length(dates), 2), 
+               dimnames = list(NULL, as.character(dates), c(1:2)))
+
+# files in this directory has different extensions, 
+# so each tile is loaded individually
+# they are also very large so it is retrieved from a external drive
+# check if the drive exists
+
+for (i in 1:2){
+  
+  if(i == 1 ) m <- "day" else m <- "night"
+  
+  for(j in 1:length(tiles)){
+    
+    file <- list.files(paste0(edrive, "modis/smooth_Ethiopia/", tiles[j]), 
+                       pattern = m, 
+                       full.names = TRUE)
+    
+    
[…]

### 2019-06-10 script/01_prepare_data.R
gy

@@ -299,16 +299,21 @@ df %<>%
 # .......................................
 # Check agronomic data ####
 
-# add plot size
+# set as numeric
 df %<>% 
   mutate(gy_gm = as.numeric(gy_gm),
          biomass_gm = as.numeric(biomass_gm),
          mean_seed_no_spike = as.numeric(mean_seed_no_spike),
          mean_tn = as.numeric(mean_tn),
-         mean_sl = as.numeric(mean_sl),
-         plot_size = ifelse(year == 2013, 0.4, 
-                            ifelse(year == 2014, 1.6,
-                                   ifelse(year == 2015, 1.2, NA))),
+         mean_sl = as.numeric(mean_sl))
+
+# normalise by plot size
+df %<>% 
+  mutate(gy_gm_raw = gy_gm,
+         plot_size = ifelse(year == 2013 & region == "Tigray", 1.6,
+                            ifelse(year == 2013 & region != "Tigray", 0.4,
+                                   ifelse(year == 2014, 1.6,
+                                          ifelse(year == 2015, 1.2, NA)))),
          gy_gm = gy_gm / plot_size)
 
 # mean tiller number

### 2019-06-10 script/04_bradley-terry_binomial.R

@@ -1,0 +1,31 @@
+library("tidyverse")
+library("magrittr")
+library("gosset")
+library("BradleyTerry2")
+
+
+
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv()
+
+output <- "output/exploring/"
+dir.create(output, recursive = TRUE, showWarnings = FALSE)
+
+#.....................................
+#.....................................
+# farmer rank ####
+
+# create grouped rankings 
+G <- to_rankings(data = df,
+                 items = "genotype",
+                 input = "farmer_rank",
+                 id = "id", 
+                 grouped.rankings = TRUE)
+
+
+object <- to_paircomp(G)
+
+
+
+

### 2019-06-10 script/0.get.data.tricot.R
GP scripts

@@ -1,0 +1,52 @@
+options(stringsAsFactors = F)
+
+library(plyr)
+library(RCurl)
+library(lme4)
+
+wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/0.get.data"
+setwd(wd)
+
+#get in row data from GITHUB
+url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/durumwheat.csv")
+csd<-read.csv(text=url)
+csd<-csd[order(csd[,"genotype"]),]
+
+#try to drop samples from OROMIA
+csd<-csd[which(csd[,"region"] %in% c("Amhara", "Tigray")),]
+
+#get a metric about getting first for each variety
+bygen<-split(csd, csd[,"genotype"])
+#set df as output
+outrough<-matrix(ncol=2, nrow=length(bygen))
+
+for (i in 1:length(bygen)){
+  tmp<-bygen[[i]]
+  
+  #get propotion of winning
+  wins<-length(which(tmp[,"farmer_rank"]==1))
+  odds<-wins/nrow(tmp)
+  
+  #get mean estimate
+  gy<-mean(tmp[,"gy_gm"], na.rm=T)
+
+  #bring everything in the output
+  outrough[i,]<-c(odds, gy)
+}
+
+colnames(outrough)<-c("win_prop", "mean_gy")
+rownames(outrough)<-unique(csd[,"genotype"])
+
+outrough<-as.data.frame(outrough)
+
+#get gy from a mixed model accounting for farmer and genotypes
+mod<-lmer(gy_gm  ~  (1|genotype) + (1|farmer_no) + (1|year), data=csd)
+idblup<-ranef(mod)
+gy<-idblup$genotype
+
+outrough<-merge(outrough, gy, by="row.names", all.x=T)
+rownames(outrough)<-outrough[,1]
+outrough<-outrough[,-1]
+colnames(outrough)[3]<-"gy_blup"
+
+save(outrough, file="rough.cs.data.estimates.Rdata")

### 2019-06-10 script/1.rrblup.R

@@ -1,0 +1,249 @@
+######################
+#this script performs GP with rrBLUP
+#####################
+options(stringsAsFactors = F)
+library(ggplot2)
+library(rrBLUP)
+library(reshape2)
+library(plyr)
+library(RCurl)
+
+wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/1.rrBLUP"
+setwd(wd)
+
+########  SNP DATA  ########
+#get in genotypes and clean them
+hmp<-read.delim("../../data/diversity.panel.svevo.positions.q10.hmp")
+snp.pos<-hmp[,1:11]
+geno<-hmp[,12:ncol(hmp)]
+rownames(geno)<-hmp[,1]
+colnames(geno)<-sub("^X", "", colnames(geno))
+
+#set a function converting genotypic data in the proper format to run rrBLUP
+convert<-function(x){
+  tmp<-as.character(x)
+  tmp[tmp=="N"]<-NA
+  tmp[tmp  %in% c("R","Y","S","W","K","M") ]<-0
+  uniqus<-sort(unique(tmp))
+  uniqus<-uniqus[!uniqus %in% c("0", NA)]
+  if(length(uniqus)>1){
+    tmp[tmp==uniqus[1]]<- 1
+    tmp[tmp==uniqus[2]]<- -1
+  }    
+  tmp[tmp==uniqus[1]]<-1
+  #overwrite
+  return(tmp)
+}
+
+#convert snps in -1,0,1 and NA
+geno2<-apply(geno, 1, convert)
+geno2<-data.frame(geno2)
+rownames(geno2)<-colnames(geno)
+geno2<-apply(geno2, 2, as.numeric)
+
+#perform imputation with rrBLUP
+imputed<-A.mat(geno2, max.missing=0.5, impute.method="mean", return.imputed=T)
+failrate<-apply(imputed$imputed,2, function(x) length(which(is.na(x))))
+todrop<-which(failrate>0)
+if(length(todrop)>1){
+  imputedcl<-imputed$imputed[,-which(failrate>0)]
+} else {
+  imputedcl<-imputed$imputed
+}
+rownames(imputedcl)<-colnames(geno)
+
+########  PHENOTYPIC DATA  ########
+#get in BLUP values and clean them
+phenos<-read.csv("../../data/Et_diversity_panel_ALL-TRAITS.BLUPS.csv")
+phenos<-phenos[!is.na(phenos[,3]),]
[…]

### 2019-06-10 manuscript/manuscript.Rmd
winprobs

@@
 
-Kauê de Sousa^1,2,^*, Matteo Dell’Acqua^3^, Jacob van Etten^2^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Dejene Mengistu^7^, Mario Enrico Pè^3^, [… add others], Svein Ø.
+Kauê de Sousa^1,2,^*, Matteo Dell’Acqua^3^, Jacob van Etten^2^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Basazen Fantahun Lakew^3^, Dejene Mengistu^7^, Mario Enrico Pè^3^, [… add others], Svein Ø.
 Solberg^1^

### 2019-06-10 script/04_bradley-terry_binomial.R

@@ -12,20 +12,99 @@ df %<>%
 output <- "output/exploring/"
 dir.create(output, recursive = TRUE, showWarnings = FALSE)
 
+
+pow <- paste0(output, "probability_of_winning.csv")
+pow %<>% 
+  read_csv()
+
 #.....................................
 #.....................................
 # farmer rank ####
 
 # create grouped rankings 
-G <- to_rankings(data = df,
+R <- to_rankings(data = df,
                  items = "genotype",
                  input = "farmer_rank",
-                 id = "id", 
-                 grouped.rankings = TRUE)
+                 id = "id")
+
+B <- rank_binomial(R, drop.null = TRUE)
+
+mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
+           id = "genotype", data = B, ref = "435ET_D")
 
 
-object <- to_paircomp(G)
+summary(mod)
+
+x <- coef(mod)
+
+names(x) <- gsub("genotype","",names(x))
+
+x <- tibble(genotype = names(x),
+            pow_rank_bt = x)
+
+pow %<>% 
+  merge(. , x, by = "genotype", all.x = TRUE) %>% 
+  as_tibble()
+
+
+#...........................
+#...........................
+# grain yield ####
+# remove NAs in grain yield
+df %>% 
+  filter(!is.na(gy_gm)) -> 
+  gy
+
+# only keep strict rankings of at least 2 distinct items
+gy %>%
+  group_by(id) %>%
+  summarise(keep = length(id)) %>%
+  mutate(keep = keep > 1) %>% 
+  filter(keep) ->
+  keep
+
+# apply the logical vector
+id <- gy$id %in% keep$id
+
+# keep selected observations
+gy <- gy[id,]
+
+YR <- to_rankings(data = gy,
+                  items = "genotype",
+                  input = "gy_gm",
+                  id = "id", 
+                  grouped.rankings = TRUE)
+
+
+BY <- rank_binomial(YR, drop.null = TRUE)
+
+
+mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
+           id = "genotype", data = BY)
 
 
 
+x <- coef(mod)
+names(x) <- gsub("genotype","",names(x))
+
+x <- tibble(genotype = names(x),
+            pow_gy_bt = x)
+
+
+pow %<>% 
+  merge(. , x, by = "genotype", all.x = TRUE) %>% 
[…]

### 2019-06-10 script/01_prepare_data.R
data

@@ -191,6 +191,7 @@ summary(as.factor(df$planting_date))
 
 
 varing <- sort(unique(df$year))
+
 # fill missing planting dates with average per year
 for(i in seq_along(varing)){
   y_i <- varing[i]
@@ -274,6 +275,24 @@ df %<>%
   distinct(genotype, .keep_all = TRUE) %>% 
   distinct(farmer_rank, .keep_all = TRUE) %>% 
   as_tibble()
+
+
+# remove those tested in less than 10 farms 
+df %>% 
+  group_by(accession) %>%  
+  count(accession) %>%
+  filter(n < 10) %>%
+  select(accession) %>%
+  t() %>%
+  as.vector() ->
+  drop
+
+
+drop <- !df$accession %in% drop
+
+df %<>% 
+  filter(drop)
+
 
 # only keep strict rankings of at least 2 distinct items
 df %>%

### 2019-06-10 script/03_model.R

@@ -1,49 +1,74 @@
+# Get log abilities from PlackettLuce and Bradley-Terry models 
+# as a proxy for genomic prediction
+
 library("tidyverse")
 library("magrittr")
 library("gosset")
 library("PlackettLuce")
+library("BradleyTerry2")
+library("svglite")
 
+#.....................................
+#.....................................
+# Read data ####
 
 df <- "data/durumwheat.csv"
 df %<>% 
   read_csv()
 
 
-output <- "output/exploring/"
+output <- "output/log-abilities/"
 dir.create(output, recursive = TRUE, showWarnings = FALSE)
 
 #.....................................
 #.....................................
 # farmer rank ####
 
-# create grouped rankings 
-G <- to_rankings(data = df,
+# create PlackettLuce rankings 
+R <- to_rankings(data = df,
                  items = "genotype",
                  input = "farmer_rank",
-                 id = "id", 
-                 grouped.rankings = TRUE)
+                 id = "id")
 
-# add explanatory vars
-# to_rankings internaly order the rankings by their id 
-# so I will keep the unique values by id and combine with the 
-# grouped_rankings
+mod <- PlackettLuce(R)
 
-df %>% 
-  arrange(id) %>% 
-  distinct(id, .keep_all = TRUE) %>% 
-  select(year, region) -> 
-  expvar
+svg(filename = paste0(output, "coeff_farmer_rank_PL.svg"),
+    width = 10,
+    height = 6.5,
+    pointsize = 12)
+plot(qvcalc(itempar(mod, log = TRUE)), las = 2)
+dev.off()
 
-G <- cbind(G, expvar)
-
-mod <- pltree(G ~ ., data = G)
-
-plot(mod)
 
 winprobs <- itempar(mod, log = TRUE)
 
-winprobs <- bind_cols(genotype = dimnames(winprobs)[[2]],
+winprobs <- bind_cols(genotype = names(winprobs),
                       pow_rank = as.vector(winprobs))
+
+
+#.....................................
+#.....................................
+# farmer rank BradleyTerry ####
+
+# get a binomial rank
+B <- rank_binomial(R, drop.null = TRUE)
+
+mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
+           id = "genotype", data = B, ref = "435ET_D")
+
+
[…]

### 2019-06-10 script/03_model.R
log-abilities

@@ -16,7 +16,6 @@ df <- "data/durumwheat.csv"
 df %<>% 
   read_csv()
 
-
 output <- "output/log-abilities/"
 dir.create(output, recursive = TRUE, showWarnings = FALSE)
 
@@ -54,7 +53,7 @@ winprobs <- bind_cols(genotype = names(winprobs),
 B <- rank_binomial(R, drop.null = TRUE)
 
 mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
-           id = "genotype", data = B, ref = "435ET_D")
+           id = "genotype", data = B)
 
 
 summary(mod)
@@ -69,6 +68,11 @@ x <- tibble(genotype = names(x),
 winprobs %<>% 
   merge(. , x, by = "genotype", all.x = TRUE) %>% 
   as_tibble()
+
+winprobs %<>% 
+  mutate(pow_rank_bt = ifelse(is.na(pow_rank_bt),
+                              0,
+                              pow_rank_bt))
 
 
 #...........................
@@ -123,7 +127,38 @@ winprobs %<>%
   as_tibble()
 
 
-write_csv(winprobs, paste0(output, "probability_of_winning.csv"))
+#.....................................
+#.....................................
+# farmer rank BradleyTerry ####
+
+# get a binomial rank
+B <- rank_binomial(YR, drop.null = TRUE)
+
+mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
+           id = "genotype", data = B)
+
+
+summary(mod)
+
+x <- coef(mod)
+
+names(x) <- gsub("genotype","",names(x))
+
+x <- tibble(genotype = names(x),
+            pow_gy_bt = x)
+
+winprobs %<>% 
+  merge(. , x, by = "genotype", all.x = TRUE) %>% 
+  as_tibble()
+
+winprobs %<>% 
+  mutate(pow_gy_bt = ifelse(is.na(pow_gy_bt),
+                            0,
+                            pow_gy_bt))
+
+
+write_csv(winprobs, paste0(output, "log-abilities.csv"))
+
 
 
 # ...................................

### 2019-06-10 script/1.rrblup.R
Add files via upload

@@ -65,14 +65,10 @@ csdata<-read.csv(text=url)
 rownames(csdata)<-csdata[,1]
 csdata<-csdata[,-1]
 
-#make sure that the data is normal!!!
-shapiro.test(csdata[,1])
-shapiro.test(csdata[,2])
-
 #add data produced locally
 load("../0.get.data/rough.cs.data.estimates.Rdata")
 #make it normally distributed
-outrough<-apply(outrough,2,log)
+#outrough<-apply(outrough,2,sqrt)
 joined.csdata<-merge(csdata, outrough, by="row.names")
 row.names(joined.csdata)<-joined.csdata[,1]
 joined.csdata<-joined.csdata[,-1]
@@ -115,8 +111,8 @@ load("GP.data.step1.Rdata")
 
 ########  PERFORM GP  ########
 #run GBLUP to perform genomic prediction
-fraction<-.5
-nruns<-100
+fraction<-.7
+nruns<-50
 tsize<-floor(length(csids)*fraction)
 
 #generate samples
@@ -138,7 +134,7 @@ for (zz in 1:length(samps)){
   traingeno<-setdiff(rownames(imputedcl), testgeno)
 
   #set training phenos and test phenos
-  traintrait<-colnames(phenos)[grep("GY.201[23]$|SPIKE..2012$|OVERALL..2012$", colnames(phenos))]
+  traintrait<-colnames(phenos)[grep("GY.201[23]$|SPIKE..201[23]$SPIKE..2012$|OVERALL..2012$", colnames(phenos))]
   testtrait<-colnames(phenos)[97:ncol(phenos)]
   
   tpheno<-phenos[rownames(phenos) %in% traingeno,traintrait]
@@ -172,6 +168,7 @@ for (zz in 1:length(samps)){
     traitout[[p]]<-accuracy
     names(traitout[[p]])<-curp
   }
+  
   traitoutdf<-do.call(cbind, traitout)
   colnames(traitoutdf)<-colnames(tpheno)
   predresults[[zz]]<-traitoutdf

### 2019-06-10 script/0.get.data.tricot.R
Add files via upload

@@ -3,22 +3,20 @@ options(stringsAsFactors = F)
 library(plyr)
 library(RCurl)
 library(lme4)
+library(bestNormalize)
 
 wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/0.get.data"
 setwd(wd)
 
-#get in row data from GITHUB
+#get in raw data from GITHUB
 url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/durumwheat.csv")
 csd<-read.csv(text=url)
 csd<-csd[order(csd[,"genotype"]),]
 
-#try to drop samples from OROMIA
-csd<-csd[which(csd[,"region"] %in% c("Amhara", "Tigray")),]
-
 #get a metric about getting first for each variety
 bygen<-split(csd, csd[,"genotype"])
 #set df as output
-outrough<-matrix(ncol=2, nrow=length(bygen))
+out.csd<-matrix(ncol=5, nrow=length(bygen))
 
 for (i in 1:length(bygen)){
   tmp<-bygen[[i]]
@@ -27,26 +25,55 @@ for (i in 1:length(bygen)){
   wins<-length(which(tmp[,"farmer_rank"]==1))
   odds<-wins/nrow(tmp)
   
-  #get mean estimate
+  #get mean estimates
   gy<-mean(tmp[,"gy_gm"], na.rm=T)
-
+  sl<-mean(tmp[,"mean_sl"], na.rm=T)
+  nsps<-mean(tmp[,"mean_seed_no_spike"], na.rm=T)
+  net<-mean(tmp[,"mean_tn"], na.rm=T)
+  
   #bring everything in the output
-  outrough[i,]<-c(odds, gy)
+  out.csd[i,]<-c(odds, gy, sl, nsps, net)
 }
 
-colnames(outrough)<-c("win_prop", "mean_gy")
-rownames(outrough)<-unique(csd[,"genotype"])
+colnames(out.csd)<-c("win_prop", "mean_gy", "sl", "nsps", "net")
+rownames(out.csd)<-unique(csd[,"genotype"])
 
-outrough<-as.data.frame(outrough)
+out.csd<-as.data.frame(out.csd)
 
 #get gy from a mixed model accounting for farmer and genotypes
-mod<-lmer(gy_gm  ~  (1|genotype) + (1|farmer_no) + (1|year), data=csd)
[…]

### 2019-06-11 script/01_prepare_data.R
data

@@ -1,6 +1,6 @@
 ####################################################
 ###### Read and clean durum wheat in Ethiopia
-# Updated 09un2019
+# Updated 11Jun2019
 ####################################################
 
 library("tidyverse")
@@ -149,84 +149,8 @@ plot(eth["GID_0"], col = "lightgrey", reset = FALSE)
 plot(lonlat, col = "blue", cex = 1,
      bg = "Steelblue1", pch = 21, add = TRUE)
 
-
-# .......................................
-# .......................................
-# Fix planting dates ####
-# planting dates 
-
-df %>% 
-  select(sowing_date) %>% 
-  mutate(sowing_date = as.character(sowing_date)) ->
-  date
-
-date %<>% 
-  mutate(sowing_date = gsub("[/]|[.]","-", sowing_date)) %>% 
-  separate(., sowing_date, into = c("x","y","z"), sep = "-") %>% 
-  mutate_all(as.integer)
-
-summary(as.factor(date$x))
-summary(as.factor(date$y))
-summary(as.factor(date$z))
-
-# planting dates must be between month 7 and 8
-date %<>% 
-  mutate(y = ifelse(y < 7 , 8, y),
-         y = ifelse(y > 8, 8, y))
-
-summary(as.factor(date$x))
-summary(as.factor(date$y))
-summary(as.factor(date$z))
-
-date %<>% 
-  mutate(planting_date = ifelse(x > 2000, paste(x, y, z, sep = "-"),
-                              ifelse(z > 2000, paste(z, y, x, sep = "-"), NA)))
-
-df %<>% 
-  mutate(planting_date = date$planting_date,
-         planting_date = as.Date(planting_date, format = "%Y-%m-%d"), 
-         planting_date = as.integer(planting_date))
-
-summary(as.factor(df$planting_date))
-
-
-varing <- sort(unique(df$year))
-
-# fill missing planting dates with average per year
-for(i in seq_along(varing)){
-  y_i <- varing[i]
-  # check which value is the closest to the mean
-  # and get the centroid
-  df %>% 
-    filter(year == y_i) %>% 
-    filter(!is.na(planting_date)) ->
-    x
-  
-  z <- mean(x$planting_date)
-  z <- unlist(x[, "planting_date"] - z)
-  z <- x[[which.min(abs(z)), "planting_date"]]
-  
-  df %<>% 
-    mutate(planting_date = ifelse(is.na(planting_date) & year == y_i,
-                                  z,
[…]

### 2019-06-11 script/00_get_data_tricot.R
110619

@@ -1,0 +1,79 @@
+options(stringsAsFactors = F)
+
+library(tidyverse)
+library(RCurl)
+library(lme4)
+library(bestNormalize)
+
+#wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/0.get.data"
+#setwd(wd)
+
+#get in raw data from GITHUB
+#url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/durumwheat.csv")
+csd<-read.csv("data/durumwheat.csv")
+
+csd<-csd[order(csd[,"genotype"]),]
+
+#get a metric about getting first for each variety
+bygen<-split(csd, csd[,"genotype"])
+#set df as output
+out.csd<-matrix(ncol=5, nrow=length(bygen))
+for (i in 1:length(bygen)){
+  tmp<-bygen[[i]]
+  
+  #get propotion of winning
+  wins<-length(which(tmp[,"farmer_rank"]==1))
+  odds<-wins/nrow(tmp)
+  
+  #get mean estimates
+  gy<-mean(tmp[,"gy_gm"], na.rm=T)
+  sl<-mean(tmp[,"mean_sl"], na.rm=T)
+  nsps<-mean(tmp[,"mean_seed_no_spike"], na.rm=T)
+  net<-mean(tmp[,"mean_tn"], na.rm=T)
+  
+  #bring everything in the output
+  out.csd[i,]<-c(odds, gy, sl, nsps, net)
+}
+
+colnames(out.csd)<-c("win_prop", "mean_gy", "sl", "nsps", "net")
+rownames(out.csd)<-unique(csd[,"genotype"])
+
+out.csd<-as.data.frame(out.csd)
+
+#get gy from a mixed model accounting for farmer and genotypes
+mod<-lmer(gy_gm  ~  (1|genotype) + (1|id) + (1|year), data=csd)
+idblup<-ranef(mod)
+gy<-idblup$genotype
+out.csd<-merge(out.csd, gy, by="row.names", all.x=T)
+rownames(out.csd)<-out.csd[,1]
+out.csd<-out.csd[,-1]
+colnames(out.csd)[ncol(out.csd)]<-"gy_blup"
+
+#now add in data from Kaue's models
+#url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/log-abilities/log-abilities.csv")
[…]

### 2019-06-11 script/1.rrblup.R

@@ -8,12 +8,12 @@ library(reshape2)
 library(plyr)
 library(RCurl)
 
-wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/1.rrBLUP"
-setwd(wd)
+#wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/1.rrBLUP"
+#setwd(wd)
 
 ########  SNP DATA  ########
 #get in genotypes and clean them
-hmp<-read.delim("../../data/diversity.panel.svevo.positions.q10.hmp")
+hmp<-read.delim("data/diversity.panel.svevo.positions.q10.hmp")
 snp.pos<-hmp[,1:11]
 geno<-hmp[,12:ncol(hmp)]
 rownames(geno)<-hmp[,1]
@@ -54,13 +54,13 @@ rownames(imputedcl)<-colnames(geno)
 
 ########  PHENOTYPIC DATA  ########
 #get in BLUP values and clean them
-phenos<-read.csv("../../data/Et_diversity_panel_ALL-TRAITS.BLUPS.csv")
+phenos<-read.csv("data/Et_diversity_panel_ALL-TRAITS.BLUPS.csv")
 phenos<-phenos[!is.na(phenos[,3]),]
 rownames(phenos)<-phenos[,3]
 phenos<-phenos[,6:ncol(phenos)]
 
 #load in the crowdsourcing data produced by Kaue
-url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/exploring/probability_of_winning.csv")
+url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/log-abilities/log-abilities.csv")
 csdata<-read.csv(text=url)
 rownames(csdata)<-csdata[,1]
 csdata<-csdata[,-1]

### 2019-06-11 script/helper_01_check_network.R

@@ -10,40 +10,10 @@ df <- "data/durumwheat.csv"
 df %<>%
   read_csv()
 
-
-# not genotyped varieties
-out <- "data/raw/not_in_genotyping.csv"
-out %<>% 
-  read_csv() %>% 
-  select(accession)
-
-vars <- bind_cols(accession = unlist(df[paste0("variety_",letters[1:4])]),
-                  region = rep(df$region, 4))
-
-
-vars %<>%
-  group_by(region) %>% 
-  distinct(accession)
-
-
-vars %<>% 
-  bind_cols(. , out = vars$accession %in% out$accession)
-
-vars %>%  
-  filter(out == TRUE) ->
-  out
-
-
-vars %>% 
-  filter(out == FALSE) ->
-  inn
-
-
-R <- to_rankings(items = df[paste0("variety_",letters[1:4])],
-                 input = df[paste0("rank_variety_",letters[1:4])])
-
-
-R <- R[, dimnames(R)[[2]] %in% inn$accession]
+R <- to_rankings(data = df,
+                 items = "genotype",
+                 input = "farmer_rank",
+                 id = "id")
 
 
 adj <- adjacency(R)
@@ -56,8 +26,10 @@ dimnames(adj) <- list(dimnames(R)[[2]], dimnames(R)[[2]])
 
 adj <- btdata(adj, return_graph = TRUE)
 
-plot.igraph(adj$graph, vertex.size = 10, edge.arrow.size = 0.1) 
 
+svg(filename = "output/network.svg", width = 12, height = 12, pointsize = 20)
+plot.igraph(adj$graph, vertex.size = 5, edge.arrow.size = 0.1) 
+dev.off()
 
 mod <- PlackettLuce(R)
 summary(mod)

### 2019-06-13 manuscript/manuscript.Rmd
13062019

--- 2019-05-31 manuscript/manuscript.Rmd
@@
 
-Kauê de Sousa^1,2,^*, Matteo Dell’Acqua^3^, Jacob van Etten^2^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Dejene Mengistu^7^, Mario Enrico Pè^3^, [… add others], Svein Ø.
-Solberg^1^
-^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
-^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
-^3^ Scuola Superiore Sant’Anna, Institute of Life Sciences, 56124 Pisa, Italy  
-^4^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
-^5^ College of Agriculture and Life Sciences, Cornell University, 14853 Ithaca, USA  
-^6^ Agricultural Research Service, United States Department of Agriculture, 148532901, Ithaca, USA  
+Kauê de Sousa^1,2^, Jacob van Etten^2^, Carlo Fadda^3^, Yosef Gebrehawaryat^3^, Jean-Luc Jannink^4,5^, Basazen Fantahun Lakew^6^, Dejene Mengistu^7^, Svein Ø.
+Solberg^1^, Mario Enrico Pè^6^, Jesse Poland^8^, [… add others], Melfa and Workaye Farmer Community, Matteo Dell’Acqua^6[*]^
+^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway ^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
+^3^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
+^4^ College of Agriculture and Life Sciences, Cornell University, 14853 Ithaca, USA  
+^5^ Agricultural Research Service, United States Department of Agriculture, 148532901, Ithaca, USA  
+^6^ Scuola Superiore Sant’Anna, Institute of Life Sciences, 56124 Pisa, Italy  
 ^7^ Bioversity International, Nepal Office, 3055 Kathmandu, Nepal 
+^8^ Kansas State University, Kansas, USA
+^[*]^To whom correspondence must be addressed <m.dellacqua@santannapisa.it>
 

### 2019-06-13 script/04_bradley-terry_binomial.R

@@ -1,110 +1,184 @@
+# Bradley-Terry model with player specific abilities
+# Kaue de Sousa
+# INN University / Bioversity International
+# Updated 12Jun2019
+
+# Load required packages 
 library("tidyverse")
 library("magrittr")
 library("gosset")
 library("BradleyTerry2")
+library("caret")
 
 
 
-df <- "data/durumwheat.csv"
-df %<>% 
-  read_csv()
+pairwise_contests <- function(object) {
+  
+  if(class(object) == "rankings") {
+    object <- to_paircomp(object)
+  }
+  
+  if(class(object) == "grouped_rankings") {
+    object <- to_paircomp(object)
+  }
+  
+  if(class(object) != "paircomp") {
+    stop("Unknown class for object")
+  }
+  
+  P <- object
+  
+  P <- as.matrix(P)
+  
+  contests <- apply(P, 1, function(X) {
+    
+    X <- X[!is.na(X)]
+    
+    players <- as.matrix(do.call(rbind, strsplit(names(X), ":")))
+    
+    X <- as.vector(ifelse(X == -1, 2, 1))
+    
+    w <- cbind(1:length(X), X)
+    
+    winner <- players[w]
+    
+    X <- ifelse(X == 2, 1, 2)
+    
+    l <- cbind(1:length(X), X)
+    
+    loser <- players[l]
+    
+    return(cbind(winner, loser))
+  })
+  
+  
+  contests <- do.call(rbind, contests)
+  
+  if (any(contests[,1] == contests[,2])) {
+    warning("winner and loser has equal values")
+  }
+  
+  
+  contests <- data.frame(contests, stringsAsFactors = FALSE)
+  
+  items <- unique(unlist(contests[,1:2]))
+  
+  contests$winner <- factor(contests$winner, levels = items)
+  contests$loser <- factor(contests$loser, levels = items)
+  
+  return(contests)
+  
+}
 
-output <- "output/exploring/"
+# create the output path for this analysis
+output <- "output/bradley-terry_gene/"
 dir.create(output, recursive = TRUE, showWarnings = FALSE)
 
-
-pow <- paste0(output, "probability_of_winning.csv")
-pow %<>% 
+# read the data
+df <- "data/durumwheat.csv"
+df %<>% 
   read_csv()
 
 #.....................................
 #.....................................
 # farmer rank ####
 
-# create grouped rankings 
+# create the contests object 
 R <- to_rankings(data = df,
[…]

### 2019-06-13 script/helper_01_check_network.R

@@ -10,6 +10,9 @@ df <- "data/durumwheat.csv"
 df %<>%
   read_csv()
 
+
+table(df$accession, df$region)
+
 R <- to_rankings(data = df,
                  items = "genotype",
                  input = "farmer_rank",

### 2019-06-17 manuscript/manuscript.Rmd
data

@@
 Solberg^1^, Mario Enrico Pè^6^, Jesse Poland^8^, [… add others], Melfa and Workaye Farmer Community, Matteo Dell’Acqua^6[*]^
-^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway ^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
+^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
+^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
 ^3^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
@@
 ^8^ Kansas State University, Kansas, USA  
-^[*]^To whom correspondence must be addressed <m.dellacqua@santannapisa.it>
+^[*]^Correspondence should be addressed to MDA (email: <m.dellacqua@santannapisa.it>)
 
@@
 
-* Is genomic selection with citizen science a viable approach?
-* Does the genetic information provide sufficient gain in the predictions for variety selection (given the cost)?
-* Can the combination of genomic and citizen science provide new insights in which traits farmers prioritise or which areas of the genome play a role?
-* Does genomic prediction using phenotypic data produced on station (agronomic traits, farmer preferences) improve accuracy in the field as measured in the tricot trial?
+* Can farmers' evaluations better predict varietal performance associated with genes?
+* Can we use crowdsourcing to improve GxE estimates?
+* What is the ability of local varieties to outperform varieties from other zones?
 
@@
 
-* Crowdsourcing citizen science showed the ability to register the climate effects as they occurr in farms, with potential to be used in plant breeding programs and risk assessments for farm adaptation[@VanEtten2019].
+* Crowdsourcing citizen science showed the ability to register the climate effects as they occur in farms, with potential to be used in plant breeding programs and risk assessments for farm adaptation[@VanEtten2019].
 

### 2019-06-17 script/01_prepare_data.R

@@ -55,8 +55,9 @@ df %<>%
          lat = ifelse(is.na(lon), NA, lat))
 
 # Download country border of ethiopia
+eth <-
 raster::getData("GADM", country = "ETH", path = "data", level = 0) %>% 
-  st_as_sf() -> eth
+  st_as_sf()
 
 # plot(eth["GID_0"], col = "lightgrey")
 # 

### 2019-06-17 script/helper_04_genotype_locations.R

@@ -1,0 +1,29 @@
+
+
+
+
+library("tidyverse")
+library("magrittr")
+library("janitor")
+library("sf")
+
+pass <- "data/passport_data_durumwheat.csv"
+pass %<>% 
+  read_csv() 
+
+
+eth <-
+  raster::getData("GADM", country = "ETH", path = "data", level = 1) %>% 
+  st_as_sf()
+
+pass %>%
+  select(lon, lat) %>%
+  filter(!is.na(lon)) %>%
+  as.matrix() %>%
+  st_multipoint() %>%
+  st_sfc(crs = st_crs(eth)) ->
+  lonlat
+
+plot(eth["GID_1"], col = "#F2F2F2", reset = FALSE)
+plot(lonlat, col = "blue", cex = 1,
+     bg = "Steelblue1", pch = 21, add = TRUE)

### 2019-06-25 script/02_add_climate_indices.R
environmental_indices

@@ -4,8 +4,7 @@
 library("tidyverse")
 library("magrittr")
 library("ClimMobTools")
-library("raster")
-
+library("caret")
 
 #.......................................
 #.......................................
@@ -19,132 +18,72 @@ df %<>%
 
 n <- nrow(df)
 
-# define some parameters
+load("data/modis.rda")
 
-# the external drive with MODIS and CHIRPS data
-edrive <- "E:/rasters/"
+load("data/chirps.rda")
 
-my_proj <- "+proj=longlat +datum=WGS84"
+df
 
-ext <- raster::extent(68, 86, 9, 33)
 
-modis_factor <- 0.02
+# Average dates for booting, flowering and maturity
+# c(60,75,135)
+spans <- as.matrix(cbind(period = c("veg", "rep", "gra", "sow2rep", "sow2gra"),
+                   ts = c(60, 30, 45, 90, 135),
+                   sumtosow = c(0, 60, 90, 0, 0)))
 
-k_to_c <- 273.15
 
-tiles <- c("h21v07","h21v08")
-
-years <- c(2012:2016)
-
-dates <- NULL
-
-# generate a sequency of 8-days from 2013 to 2015
-for(i in c(2012:2015)){
-  dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), 
-                          as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
-}
-
-dates <- as.Date(c(dates, seq(as.Date("2016-01-01", "%Y-%m-%d"), 
-                              as.Date("2016-12-26", "%Y-%m-%d"), 8)), 
-                 origin = "1970-01-01")
-
-# and a vector with full dates (1 days)
-full_dates <- seq(dates[1], dates[length(dates)], 1)
-
-#.......................................
-#.......................................
-# Add MODIS data ####
-
-# create an array with 2 layers
-# modis day and night
-modis <- array(NA, dim = c(n, length(dates), 2), 
-               dimnames = list(NULL, as.character(dates), c(1:2)))
-
-# files in this directory has different extensions, 
-# so each tile is loaded individually
-# they are also very large so it is retrieved from a external drive
-# check if the drive exists
-
-for (i in 1:2){
+# ............................................
+# ............................................
+# rainfall indices ####
+rain <- NULL
[…]

### 2019-06-25 script/05_model.R

@@ -1,0 +1,54 @@
+# 
+library("tidyverse")
+library("magrittr")
+library("PlackettLuce")
+library("gosset")
+
+
+#...................................
+#...................................
+# read data 
+list.files("data")
+
+# farmer rankings
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv()
+
+
+
+# additive matrix from genes
+load("data/additive.matrix.cs.rda")
+
+
+# normal prior object from additive matrix
+prior <- list(mu = rep(0, ncol(additivemat)),
+              Sigma = additivemat)
+
+
+#.....................................
+#.....................................
+# farmer rank ####
+
+# create PlackettLuce rankings 
+R <- to_rankings(data = df,
+                 items = "genotype",
+                 input = "farmer_rank",
+                 id = "id", 
+                 grouped.rankings = TRUE)
+
+
+mod <- PlackettLuce(R, npseudo = 0)
+
+modprior <- PlackettLuce(R, normal = prior, gamma = TRUE)
+
+
+AIC(mod)
+AIC(modprior)
+
+mod$adherence
+
+modprior$adherence
+
+modprior$normal
+

### 2019-06-25 script/helper_04_genotype_locations.R

@@ -17,8 +17,11 @@ eth <-
   st_as_sf()
 
 pass %>%
-  select(lon, lat) %>%
-  filter(!is.na(lon)) %>%
+  dplyr::select(lon, lat) %>%
+  filter(!is.na(lon)) ->
+  xy
+  
+xy %>% 
   as.matrix() %>%
   st_multipoint() %>%
   st_sfc(crs = st_crs(eth)) ->
@@ -27,3 +30,5 @@ pass %>%
 plot(eth["GID_1"], col = "#F2F2F2", reset = FALSE)
 plot(lonlat, col = "blue", cex = 1,
      bg = "Steelblue1", pch = 21, add = TRUE)
+
+

### 2019-06-25 script/helper_05_match_farmers_ids.R

@@ -1,0 +1,122 @@
+# Find matches between citizen science dataset
+# and a farmer survey in Ethiopia
+#
+# Kauê de Sousa
+
+library(tidyverse)
+library(magrittr)
+
+# read Martina's household data with her ids
+hh <- "data/HH_id2019_Ethiopia.csv"
+hh %<>% 
+  read_csv()
+
+# also the file with kebeles ids
+kb <- "data/HH_id2019_Ethiopia_kebeles.csv"
+kb %<>% 
+  read_csv() %>% 
+  select(-district_id)
+
+# join the two datasets
+hh %<>% 
+  inner_join(., kb, by = "kebele_id")
+
+# read the citizen science data
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv() %>% 
+  select(id, farmer, region, kebele) %>% 
+  distinct(id, .keep_all = TRUE)
+
+# get a vector with unique names for kebeles
+kebele <- sort(unique(df$kebele))
+
+kebele[!kebele %in% sort(unique(hh$kebele))]
+
+# fix some issues in kebeles names
+# first colunm is the names that appear in the citizen science data
+# the second is in Martina's survey
+repl <- matrix(c("Adi kunti","Adi Kuenti",
+                 "Koka","Koca",
+                 "Burtilik","Birtilik",
+                 "Medagolat","Meda golat" ,
+                 "Mysedri","May sedri",
+                 "Tility","Tlity",
+                 "Weketa","Wekete",
+                 "Werkaye","Workaye",
+                 "Wole","Wole Amba",
+                 "Weyn amba","Weyra Amba"), 
+               nrow = 10, ncol = 2, byrow = TRUE)
+
+# replace names in the citizen science dataset
+for(i in seq_along(repl[,1])) {
+  df$kebele <- ifelse(df$kebele == repl[i,1],
+                      repl[i,2],
+                      df$kebele)
+}
+
+# get a new vector with kebele names
+kebele <- sort(unique(df$kebele))
+
+# check the kebeles in Martina's data that appear in the 
+# citizen science dataset
+kebele <- kebele[kebele %in% sort(unique(hh$kebele))]
+
+
+
+# keep only kebeles that occurs in both datasets
+keep <- df$kebele %in% kebele
+df <- df[keep, ]
+
+keep <- hh$kebele %in% kebele
+hh <- hh[keep,]
+
+# empty vector to match with survey dataset
+df$hhid <- NA
[…]

### 2019-06-25 script/helper_06_compile_chirps_modis.R

@@ -1,0 +1,150 @@
+# Add environmental variables using planting dates 
+# and lon lat
+
+library("tidyverse")
+library("magrittr")
+library("ClimMobTools")
+library("raster")
+
+
+#.......................................
+#.......................................
+# Read data ####
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv()
+
+df %<>%
+  distinct(id, .keep_all = TRUE)
+
+n <- nrow(df)
+
+# define some parameters
+
+# the external drive with MODIS and CHIRPS data
+edrive <- "E:/rasters/"
+
+my_proj <- "+proj=longlat +datum=WGS84"
+
+ext <- raster::extent(68, 86, 9, 33)
+
+modis_factor <- 0.02
+
+k_to_c <- 273.15
+
+tiles <- c("h21v07","h21v08")
+
+years <- c(2012:2016)
+
+dates <- NULL
+
+# generate a sequency of 8-days from 2013 to 2015
+for(i in c(2012:2015)){
+  dates <- c(dates, c(seq(as.Date(paste0(i, "-01-01"), "%Y-%m-%d"), 
+                          as.Date(paste0(i, "-12-31"), "%Y-%m-%d"), 8)))
+}
+
+dates <- as.Date(c(dates, seq(as.Date("2016-01-01", "%Y-%m-%d"), 
+                              as.Date("2016-12-26", "%Y-%m-%d"), 8)), 
+                 origin = "1970-01-01")
+
+# and a vector with full dates (1 days)
+full_dates <- seq(dates[1], dates[length(dates)], 1)
+
+#.......................................
+#.......................................
+# Add MODIS data ####
+
+# create an array with 2 layers
+# modis day and night
+modis <- array(NA, dim = c(n, length(dates), 2), 
+               dimnames = list(NULL, as.character(dates), c(1:2)))
+
+# files in this directory has different extensions, 
+# so each tile is loaded individually
+# they are also very large so it is retrieved from a external drive
+# check if the drive exists
+
+for (i in 1:2){
+  
+  if(i == 1 ) m <- "day" else m <- "night"
+  
+  for(j in 1:length(tiles)){
+    
+    file <- list.files(paste0(edrive, "modis/smooth_Ethiopia/", tiles[j]), 
+                       pattern = m, 
+                       full.names = TRUE)
+    
+    
[…]

### 2019-06-28 manuscript/manuscript.Rmd
ms

@@
 
-Kauê de Sousa^1,2^, Jacob van Etten^2^, Carlo Fadda^3^, Yosef Gebrehawaryat^3^, Jean-Luc Jannink^4,5^, Basazen Fantahun Lakew^6^, Dejene Mengistu^7^, Svein Ø.
-Solberg^1^, Mario Enrico Pè^6^, Jesse Poland^8^, [… add others], Melfa and Workaye Farmer Community, Matteo Dell’Acqua^6[*]^
+Kauê de Sousa^1,2^, Jacob van Etten^2^, Svein Ø.
+Solberg^1^, Mario Enrico Pè^3^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Basazen Fantahun Lakew^3^, Dejene Mengistu^7^,  Jesse Poland^8^, [… add others], Matteo Dell’Acqua^3[*]^
 ^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
 ^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
-^3^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
-^4^ College of Agriculture and Life Sciences, Cornell University, 14853 Ithaca, USA  
-^5^ Agricultural Research Service, United States Department of Agriculture, 148532901, Ithaca, USA  
-^6^ Scuola Superiore Sant’Anna, Institute of Life Sciences, 56124 Pisa, Italy  
+^3^ Scuola Superiore Sant’Anna, Institute of Life Sciences, 56124 Pisa, Italy  
+^4^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
+^5^ College of Agriculture and Life Sciences, Cornell University, 14853 Ithaca, USA  
+^6^ Agricultural Research Service, United States Department of Agriculture, 148532901, Ithaca, USA  
 ^7^ Bioversity International, Nepal Office, 3055 Kathmandu, Nepal  
@@
 
-**Research questions:**
+# Research questions:
 
-* Can farmers' evaluations better predict varietal performance associated with genes?
-* Can we use crowdsourcing to improve GxE estimates?
-* What is the ability of local varieties to outperform varieties from other zones?
-
-# Introduction
-
-* Crowdsourcing citizen science showed the ability to register the climate effects as they occur in farms, with potential to be used in plant breeding programs and risk assessments for farm adaptation[@VanEtten2019].
+* Does farmers' evaluations associated with genes provide better predictions on varietal performance then station-only data?
+* Does farmer's crowdsourced evaluations associated to SNP markers improve GxE estimates?
+* Can we bring a new breeding paradigm where farmers actively participate in plant selection?
 
@@
 
-## Model approaches
+## Genomic prediction
+
+Methods from RBlup to genomic prediction with SNP markers.
+Farmers' evaluation from citizen science trials are connect to this approach by adding the probability of winning from Plackett-Luce as explanatory variables.
+
+Other explanatory variables include the metric traits from station experiments.
+
+## GxE interactions
+
+Split varieties from citizen science trails by groups from their origin (cold tolerants and low regions). This cal also be used in the genomic prediction.
 
@@
 
-# Results
+# Topics for discussion (or keep in mind)
 
-# Discussion
-
-* What is the relative costs of these approaches
+* GxE driven by under-representative samples or farm management.
 * Is data missingness non-random? Can data missingness be predicted from farmer preferences? 

### 2019-06-28 script/02_add_climate_indices.R

@@ -28,8 +28,8 @@ df
 # Average dates for booting, flowering and maturity
 # c(60,75,135)
 spans <- as.matrix(cbind(period = c("veg", "rep", "gra", "sow2rep", "sow2gra"),
-                   ts = c(60, 30, 45, 90, 135),
-                   sumtosow = c(0, 60, 90, 0, 0)))
+                   ts = c(65, 30, 45, 95, 140),
+                   sumtosow = c(0, 65, 95, 0, 0)))
 
 
 # ............................................
@@ -74,8 +74,8 @@ temp
 # combine temperature and rainfall indices
 indices <- bind_cols(rain, temp)
 
-indices
-
+indices %<>% 
+  mutate(year = df$year)
 
 # remove variables with low variance
 out <- nearZeroVar(indices)

### 2019-06-28 script/05_model.R

@@ -3,6 +3,9 @@ library("tidyverse")
 library("magrittr")
 library("PlackettLuce")
 library("gosset")
+library("foreach")
+library("doParallel")
+library("abind")
 
 
 #...................................
@@ -15,40 +18,222 @@ df <- "data/durumwheat.csv"
 df %<>% 
   read_csv()
 
-
-
 # additive matrix from genes
 load("data/additive.matrix.cs.rda")
 
 
-# normal prior object from additive matrix
-prior <- list(mu = rep(0, ncol(additivemat)),
-              Sigma = additivemat)
-
+# environmental indices
+ind <- "data/environmental_indices.csv"
+ind %<>% 
+  read.csv()
 
 #.....................................
 #.....................................
-# farmer rank ####
-
 # create PlackettLuce rankings 
-R <- to_rankings(data = df,
+G <- to_rankings(data = df,
                  items = "genotype",
                  input = "farmer_rank",
                  id = "id", 
                  grouped.rankings = TRUE)
 
 
-mod <- PlackettLuce(R, npseudo = 0)
+mod <- PlackettLuce(G)
 
-modprior <- PlackettLuce(R, normal = prior, gamma = TRUE)
+# normal prior object from additive matrix
+prior <- list(mu = as.vector(coef(mod)),
+              Sigma = additivemat)
 
 
-AIC(mod)
-AIC(modprior)
+#.................................................
+#.................................................
+# Set parameters for forward selection ####
 
-mod$adherence
+data <- cbind(G, ind[50:65])
 
-modprior$adherence
+# add a empty variable to the model 
+data$empty_model <- 1
 
-modprior$normal
+folds <- as.integer(as.factor(ind$year))
 
+k <- max(folds)
+
+vars <- names(data)[2:ncol(data)]
+
+minsize <- round((nrow(data)*0.15), -1)
+
+bonferroni <- TRUE
+
+alpha <- 0.05
+
+ncores <- 2
+
+# a list to keep the parameters from each step 
+coeffs <- list()
+
+# create cluster to do parallelisation
+cluster <- parallel::makeCluster(ncores)
+doParallel::registerDoParallel(cluster)
+
+# keep running until the model get its best performance
+while (best) {
+  
+  cat("\nForward Selection. Step ", counter, "\n Time: ", date(), "\n")
+  
+  fs <- length(exp_var)
+  
+  args <- list(data = data, 
+               k = k, 
+               folds = folds)
+  
+  args <- c(args, dots)
+  
+  i <- 1:fs
+  
[…]

### 2019-07-01 manuscript/manuscript.Rmd
text

@@
 ---
-title: "Combining genomic prediction with citizen science for crop improvement and early selection in durum wheat"
+title: "Genomic prediction associated with citizen science for early selection and improvement in durum wheat"
 bibliography: "library/tricotgenomic.bib"
@@
     reference_docx: "library/word-styles-reference-01.docx"
-# output: pdf_document
 ---
@@
 ^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica   
-^3^ Scuola Superiore Sant’Anna, Institute of Life Sciences, 56124 Pisa, Italy  
+^3^ Institute of Life Sciences, Scuola Superiore Sant’Anna, 56124 Pisa, Italy  
 ^4^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
-^5^ College of Agriculture and Life Sciences, Cornell University, 14853 Ithaca, USA  
-^6^ Agricultural Research Service, United States Department of Agriculture, 148532901, Ithaca, USA  
+^5^ College of Agriculture and Life Sciences, Cornell University, NY 14853 Ithaca, USA  
+^6^ Agricultural Research Service, United States Department of Agriculture, NY 14853 Ithaca, USA  
 ^7^ Bioversity International, Nepal Office, 3055 Kathmandu, Nepal  
-^8^ Kansas State University, Kansas, USA  
+^8^ Department of Agronomy, Kansas State University, KS 66506 Manhattan, USA  
 ^[*]^Correspondence should be addressed to MDA (email: <m.dellacqua@santannapisa.it>)
@@
 * Does farmers' evaluations associated with genes provide better predictions on varietal performance then station-only data?
-* Does farmer's crowdsourced evaluations associated to SNP markers improve GxE estimates?
-* Can we bring a new breeding paradigm where farmers actively participate in plant selection?
+* Can the GxE estimates be improved by farmers' evaluations associated to SNP markers?
+* Can the combination of genomic and citizen science provide new insights in which traits farmers prioritise or which areas of the genome play a role?
 
@@
 
-Methods from RBlup to genomic prediction with SNP markers.
-Farmers' evaluation from citizen science trials are connect to this approach by adding the probability of winning from Plackett-Luce as explanatory variables.
-
+RBlup to genomic prediction with SNP markers.
+Farmers' evaluation from citizen science trials are connected to this approach by adding the probability of winning from Plackett-Luce as explanatory variables.
 Other explanatory variables include the metric traits from station experiments.
@@
 
-Split varieties from citizen science trails by groups from their origin (cold tolerants and low regions). This cal also be used in the genomic prediction.
+Split varieties from citizen science trails by groups from their origin (cold tolerants vs others). This can also be used in the genomic prediction.
 
-Run a number of scenarios from varieties selection using the citizen science and station data to determine whether the genetic information provides any gain or not
+Run a number of scenarios from varieties selection using the citizen science and station data to determine whether the genetic information provides any gain or not.
 

### 2019-07-03 script/02_add_climate_indices.R
scripts

@@ -78,7 +78,7 @@ indices %<>%
   mutate(year = df$year)
 
 # remove variables with low variance
-out <- nearZeroVar(indices)
+out <- nearZeroVar(indices, freqCut = 95/50, uniqueCut = 50)
 
 names(indices)[out]
 

### 2019-07-03 script/05_model.R

@@ -7,6 +7,7 @@ library("foreach")
 library("doParallel")
 library("abind")
 
+source("script/helper_07_forward.R")
 
 #...................................
 #...................................
@@ -37,6 +38,8 @@ G <- to_rankings(data = df,
                  grouped.rankings = TRUE)
 
 
+P <- to_paircomp(G)
+
 mod <- PlackettLuce(G)
 
 # normal prior object from additive matrix
@@ -47,30 +50,32 @@ prior <- list(mu = as.vector(coef(mod)),
 #.................................................
 #.................................................
 # Set parameters for forward selection ####
+out <- !grepl("Rx1day|Rx5day|SU|Rtotal|SDII", names(ind))
+ind <- ind[out]
 
-data <- cbind(G, ind[50:65])
+data <- cbind(P, ind[-ncol(ind)])
 
-# add a empty variable to the model 
-data$empty_model <- 1
-
+# cross-validation parameters
 folds <- as.integer(as.factor(ind$year))
-
 k <- max(folds)
 
-vars <- names(data)[2:ncol(data)]
-
+# PlackettLuce parameters
 minsize <- round((nrow(data)*0.15), -1)
-
-bonferroni <- TRUE
-
+bonferroni <- FALSE
 alpha <- 0.05
+normal <- prior
+npseudo <- 0.5
 
-ncores <- 2
-
-# a list to keep the parameters from each step 
+# forward selection parameters 
+data$empty_model <- 1
+vars <- names(data)[2:ncol(data)]
+var_keep <- character()
 coeffs <- list()
+counter <- 1
+best <- TRUE
 
-# create cluster to do parallelisation
+# parallelisation parameters
+ncores <- 2
 cluster <- parallel::makeCluster(ncores)
 doParallel::registerDoParallel(cluster)
 
@@ -79,30 +84,29 @@ while (best) {
   
   cat("\nForward Selection. Step ", counter, "\n Time: ", date(), "\n")
   
-  fs <- length(exp_var)
+  fs <- length(vars)
   
   args <- list(data = data, 
                k = k, 
-               folds = folds)
-  
-  args <- c(args, dots)
+               folds = folds, 
+               alpha = alpha,
+               minsize = minsize)
   
   i <- 1:fs
   
   # get predictions from nodes and put in matrix
-  models <- try(foreach::foreach(i = i,
-                                 .combine = .comb,
-                                 .packages = packages) %dopar% (.forward_dopar(as.formula(
[…]

### 2019-07-03 script/helper_01_check_network.R

@@ -21,6 +21,8 @@ R <- to_rankings(data = df,
 
 adj <- adjacency(R)
 
+connectivity(adj)
+
 adj <- as.vector(adj)
 
 adj <- t(matrix(adj, nrow = ncol(R), ncol = ncol(R)))

### 2019-07-03 script/helper_07_forward.R

@@ -1,0 +1,87 @@
+# combine results from parallel
+.comb <- function(...) {
+  abind::abind(..., along = 1, force.array = TRUE)
+}
+
+# model call for parallel
+.forward_dopar <- function(formula, args){
+  
+  args <- c(formula, args)
+  
+  m <- do.call(gosset::crossvalidation, args)
+  
+  result <- m$raw$estimators
+  
+  nfold <- max(m$raw$folds)
+  
+  result  <- array(unlist(result), c(1, nfold, 6))
+  
+  return(result)
+  
+}
+
+
+# logical function for > greater 
+.is_greater <- function(x, y) {
+  x > y
+}
+
+
+# logical function for < lower
+.is_lower <- function(x, y) {
+  x < y
+}
+
+
+# Compute weighted means in cross-validation
+.mean_crossvalidation <- function(object, folds = NULL, 
+                                  mean.method = NULL, 
+                                  ...){
+  # take length of folds
+  N <- length(folds)
+  
+  if (is.null(mean.method)) {mean.method <- "stouffer"}
+  
+  # weight of imbalanced folds
+  if (mean.method == "stouffer") {
+    # take the number of folds
+    max_folds <- max(folds)
+    
+    # make a table of folds and take
+    # how many observations each fold has
+    foldsize <- table(folds)
+    
+    # take the weight of each fold
+    # first, the squared root of foldsize (observations per fold)
+    # by the total number of observation
+    wfold <- sqrt(as.vector(foldsize) / N)
+    
+    # then divide this vector by its sum
+    wfold <- wfold / sum(wfold)
+    # wfold <- (max_folds * wfold) / sum(wfold)
+    
+    # then we multiply the input values by the
+    # weight of each fold
+    stouffer <- object * wfold
+    
+    # sum these values and that is the stouffer mean
+    mean <- sum(stouffer)
+  }
+  
+  if (mean.method == "foldsize") {
+    # make a table of folds and take
+    # the number of observations per fold
+    foldsize <- as.vector(table(folds))
+    
+    # fold size mean is the product of multiplication of object values by 
+    # its number of observations then divided by the total number of observations
+    mean <- sum(object * foldsize, na.rm = TRUE) / sum(foldsize, na.rm = TRUE)
+  }
+  
+  if (mean.method == "equal") {
+    mean <- mean(object, na.rm = TRUE)
+  }
+  
+  return(mean)
+  
+}

### 2019-07-25 manuscript/manuscript.Rmd
plmodel

@@
 
-# Research questions:
-
-* Does farmers' evaluations associated with genes provide better predictions on varietal performance then station-only data?
-* Can the GxE estimates be improved by farmers' evaluations associated to SNP markers?
-* Can the combination of genomic and citizen science provide new insights in which traits farmers prioritise or which areas of the genome play a role?
+# Introduction 
+- wheat is the second largest ceral grain cultivated worldwide
+- durum wheat is mainly used for pasta production 
+- is a cultivar evolved from domesticated emmer wheat 
+- most of its genetic diversity is found in Ethiopia 
+- even though most farmers are exposed to a small set of Mediterranean modern varieties
+-
 

### 2019-07-25 script/00_get_data_tricot.R

@@ -1,79 +1,0 @@
-options(stringsAsFactors = F)
-
-library(tidyverse)
-library(RCurl)
-library(lme4)
-library(bestNormalize)
-
-#wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/0.get.data"
-#setwd(wd)
-
-#get in raw data from GITHUB
-#url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/durumwheat.csv")
-csd<-read.csv("data/durumwheat.csv")
-
-csd<-csd[order(csd[,"genotype"]),]
-
-#get a metric about getting first for each variety
-bygen<-split(csd, csd[,"genotype"])
-#set df as output
-out.csd<-matrix(ncol=5, nrow=length(bygen))
-for (i in 1:length(bygen)){
-  tmp<-bygen[[i]]
-  
-  #get propotion of winning
-  wins<-length(which(tmp[,"farmer_rank"]==1))
-  odds<-wins/nrow(tmp)
-  
-  #get mean estimates
-  gy<-mean(tmp[,"gy_gm"], na.rm=T)
-  sl<-mean(tmp[,"mean_sl"], na.rm=T)
-  nsps<-mean(tmp[,"mean_seed_no_spike"], na.rm=T)
-  net<-mean(tmp[,"mean_tn"], na.rm=T)
-  
-  #bring everything in the output
-  out.csd[i,]<-c(odds, gy, sl, nsps, net)
-}
-
-colnames(out.csd)<-c("win_prop", "mean_gy", "sl", "nsps", "net")
-rownames(out.csd)<-unique(csd[,"genotype"])
-
-out.csd<-as.data.frame(out.csd)
-
-#get gy from a mixed model accounting for farmer and genotypes
-mod<-lmer(gy_gm  ~  (1|genotype) + (1|id) + (1|year), data=csd)
-idblup<-ranef(mod)
-gy<-idblup$genotype
-out.csd<-merge(out.csd, gy, by="row.names", all.x=T)
-rownames(out.csd)<-out.csd[,1]
-out.csd<-out.csd[,-1]
-colnames(out.csd)[ncol(out.csd)]<-"gy_blup"
-
-#now add in data from Kaue's models
-#url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/log-abilities/log-abilities.csv")
[…]

### 2019-07-25 script/04_plmodel.R

@@ -1,0 +1,199 @@
+# Model scenarios with explanatory variables 
+# to predict variety performance using PlackettLuce models
+
+library("tidyverse")
+library("magrittr")
+library("PlackettLuce")
+library("gosset")
+library("foreach")
+library("doParallel")
+library("abind")
+source("script/helper_07_forward.R")
+
+#.................................................
+#.................................................
+# Data ####
+list.files("data")
+
+# farmer rankings
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv()
+
+
+# additive matrix from gene markers
+load("data/additive.matrix.cs.rda")
+
+
+# environmental indices
+ind <- "data/environmental_indices.csv"
+ind %<>% 
+  read.csv()
+
+#.................................................
+#.................................................
+# PlackettLuce rankings ####
+G <- to_rankings(data = df,
+                 items = "genotype",
+                 input = "farmer_rank",
+                 id = "id", 
+                 grouped.rankings = TRUE)
+
+# normal prior coefficients
+mod <- PlackettLuce(G)
+prior <- list(mu = as.vector(coef(mod)),
+              Sigma = additivemat)
+
+items <- names(coef(mod))
+
+
+#.................................................
+#.................................................
+# Forward selection ####
+
+output <- "output/plmodels/"
+dir.create(output,
+           showWarnings = FALSE,
+           recursive = TRUE)
+
+# cross validation parameters
+n <- length(G)
+minsize <- floor(n*0.2)
+gamma <- TRUE
+alpha <- 0.01
+npseudo <- 5
+
+set.seed(123)
+k <- n
+folds <- sample(rep(1:k, times = ceiling(n / k), length.out = n))
+
+keep <- grepl("DT_|NT_", names(ind))
+
+data <- cbind(G, empty_model = 0, ind[keep])
+
+# cvnull <- crossvalidation(G ~ empty_model,
+#                           data = data,
+#                           k = k,
+#                           folds = folds,
+#                           npseudo = npseudo,
+#                           alpha = alpha
+#                           
+# args <- list(data = data, 
+#              k = k, 
+#              folds = folds,
+#              minsize = minsize,
+#              gamma = gamma,
+#              alpha = alpha,
+#              npseudo = npseudo)
+# 
+# # create cluster to do parallelisation
+# ncores <- detectCores() -1
+# cluster <- parallel::makeCluster(ncores)
[…]

### 2019-07-25 script/05_model.R

@@ -1,140 +1,0 @@
-library("tidyverse")
-library("magrittr")
-library("PlackettLuce")
-library("gosset")
-library("foreach")
-library("doParallel")
-library("abind")
-
-source("script/helper_07_forward.R")
-
-#...................................
-#...................................
-# read data 
-list.files("data")
-
-# farmer rankings
-df <- "data/durumwheat.csv"
-df %<>% 
-  read_csv()
-
-
-# additive matrix from genes
-load("data/additive.matrix.cs.rda")
-
-
-# environmental indices
-ind <- "data/environmental_indices.csv"
-ind %<>% 
-  read.csv()
-
-#.....................................
-#.....................................
-# create PlackettLuce rankings ####
-G <- to_rankings(data = df,
-                 items = "genotype",
-                 input = "farmer_rank",
-                 id = "id", 
-                 grouped.rankings = TRUE)
-
-mod <- PlackettLuce(G)
-
-# normal prior object from additive matrix
-prior <- list(mu = as.vector(coef(mod)),
-              Sigma = additivemat)
-
-#.................................................
-#.................................................
-# Set parameters for forward selection ####
-#out <- !grepl("Rx1day|Rx5day|SU|Rtotal|SDII", names(ind))
-#ind <- ind[out]
-keep <- grepl("DT_|NT_|lon|lat|year", names(ind))
-
-data <- cbind(G, empty_model = 0, ind[keep])
-
-n <- nrow(data)
-
-minsize <- floor(n*0.15)
-
-null <- pltree(G ~ empty_model,
-               data = data,
-               npseudo = 5,
-               normal = prior)
-
-design <- pltree(G ~ lon + lat + year,
-               data = data,
-               npseudo = 5,
-               minsize = minsize)
-
-clim <- pltree(G ~ minDT_veg + minNT_sow2gra,
-               data = data,
-               npseudo = 5,
-               gamma = TRUE,
-               minsize = minsize)
-
-climgen <- pltree(G ~ minDT_veg + minNT_sow2gra,
-                  data = data,
-                  npseudo = 5,
-                  gamma = TRUE,
-                  normal = prior,
-                  minsize = minsize)
-
-predict(null)
-predict(design)
-predict(clim)
-predict(climgen)
-
-pseudoR2(null)
-pseudoR2(design)
-pseudoR2(clim)
-pseudoR2(climgen)
-
-
-
-# cross validation 
-set.seed(123)
-k <- n
-folds <- sample(rep(1:k, times = ceiling(n / k), length.out = n))
-
[…]

### 2019-07-25 script/helper_07_forward.R

@@ -1,6 +1,11 @@
+library("abind")
+library("gosset")
+library("parallel")
+library("doParallel")
+
 # combine results from parallel
 .comb <- function(...) {
-  abind::abind(..., along = 1, force.array = TRUE)
+  abind::abind(..., along = 2, force.array = TRUE)
 }
 
 # model call for parallel
@@ -10,12 +15,8 @@
   
   m <- do.call(gosset::crossvalidation, args)
   
-  result <- m$raw$estimators
-  
-  nfold <- max(m$raw$folds)
-  
-  result  <- array(unlist(result), c(1, nfold, 6))
-  
+  result <- m$raw$estimators$deviance
+
   return(result)
   
 }

### 2019-07-25 script/helper_03_genotype_locations.R
genotype_locations

@@ -1,34 +1,21 @@
-
-
-
-
+# create a SpatialPoints object from accessions locations
 library("tidyverse")
 library("magrittr")
-library("janitor")
-library("sf")
+library("rgeos")
+library("rgdal")
 
 pass <- "data/passport_data_durumwheat.csv"
 pass %<>% 
   read_csv() 
 
+pass %<>% 
+  filter(!is.na(lon)) %>% 
+  select(genotype, lon, lat)
 
-eth <-
-  raster::getData("GADM", country = "ETH", path = "data", level = 1) %>% 
-  st_as_sf()
+proj <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
 
-pass %>%
-  dplyr::select(lon, lat) %>%
-  filter(!is.na(lon)) ->
-  xy
-  
-xy %>% 
-  as.matrix() %>%
-  st_multipoint() %>%
-  st_sfc(crs = st_crs(eth)) ->
-  lonlat
+p <- SpatialPointsDataFrame(pass[c("lon","lat")], pass, 
+                            proj4string = CRS(proj))
 
-plot(eth["GID_1"], col = "#F2F2F2", reset = FALSE)
-plot(lonlat, col = "blue", cex = 1,
-     bg = "Steelblue1", pch = 21, add = TRUE)
 
 

### 2019-08-28 manuscript/manuscript.Rmd
20190828

@@
 - wheat is the second largest ceral grain cultivated worldwide
-- durum wheat is mainly used for pasta production 
-- is a cultivar evolved from domesticated emmer wheat 
-- most of its genetic diversity is found in Ethiopia 
+- most of durum wheat's genetic diversity is found in Ethiopia 
 - even though most farmers are exposed to a small set of Mediterranean modern varieties
@@
 
+# Part 1
+
+```{r, fig.margin=TRUE, echo=FALSE, fig.cap="Fig. 1. Research method."}
+knitr::include_graphics("../output/scheme/Fig1.png")
+```
+
+# Part 2
+
+Farmer evaluation traits can predict metric traits fairly well, depending on their derivation.
+Overall, spike, and GY observed in first year have a similar performance in predicting GY in the second year (Fig.
+2A).
+Farmers scores (both overall and spike) well predict TGW in 2013.
+There is a gender interaction, especially in overall, as arguably men and women have a different interpretation of the overall appreciation of varieties.
+When men and women criteria are contrasting, the prediction power decreases (see overall, overall.f and overall.m over GY in 2013).
+When they are aligned, the prediction power increases (see overall, overall.f and overall.m over DF in 2013; farmers are very worried about late maturation of varieties).
+
+```{r echo=FALSE, fig.cap="Fig. 2. R", fig.height=1, fig.margin=TRUE, fig.width=1}
+knitr::include_graphics("../output/gp/Fig2.png")
+```
+
+# Part 3
+
+```{r, fig.margin=TRUE, echo=FALSE, fig.cap="Fig.
+3.
+Expected yield gain of adopting variety recommendations based on (L) location model, (E) environmental model and (GxE) genotype by environmental model.", fig.height=5, fig.width=5}
+knitr::include_graphics("../output/yield_gain/yield_gain.png")
+```
+
 # Methods
 
-## Genomic prediction
+# References
 
-RBlup to genomic prediction with SNP markers.
-Farmers' evaluation from citizen science trials are connected to this approach by adding the probability of winning from Plackett-Luce as explanatory variables.
-Other explanatory variables include the metric traits from station experiments.
-
-## GxE interactions
-
-Split varieties from citizen science trails by groups from their origin (cold tolerants vs others). This can also be used in the genomic prediction.
-
-Run a number of scenarios from varieties selection using the citizen science and station data to determine whether the genetic information provides any gain or not.
-
-**Scenario 1**
-
-Baseline: tricot without genomic prediction, blocked cross-validation with seasons as blocks
-
-* Plackett-Luce model
-* Bradley-Terry model. Improvement: same exercise, but adding genomic data as explanatory data
-* Plackett-Luce model with covariance matrix derived from genomic data
-* Bradley-Terry model with genomic random variables in a GLM framework(?)
-
-**Scenario 2**
-
-Baseline: farmer preferences from station rating exercise, predicting field performance in tricot perhaps stratifying by environment and using the two station sites separately for their respective most similar environments (to circumvent GxE).
-
-Predict rating data with an appropriate GLM model for ordinal data (the rates should not be assumed to be equidistant, the difference between rates 1 and 2 may be different from the rates between 4 and 5).
-
-Improvement: same farmer preferences + genomic data, letting the ordering be influenced by genomic relatedness
-The exercise could use rank correlation, Kendall´s tau as criterion, but could also make a real “selection” (say top 50%) of the accessions and determine the gain as the % that indeed falls in the top 50% in the tricot trial.
-A p value for this exercise could be determined by doing 10,000 random selections of 50% to get a null distribution.
-
-# Topics for discussion (or keep in mind)
-
-* GxE driven by under-representative samples or farm management.
-* Is data missingness non-random? Can data missingness be predicted from farmer preferences? 
-* Can we say something about which genomic regions play a role?
-
-# References
+# Display items

### 2019-08-28 script/01_prepare_data.R

@@ -1,7 +1,8 @@
-####################################################
-###### Read and clean durum wheat in Ethiopia
-# Updated 11Jun2019
-####################################################
+# ..........................................
+# ..........................................
+# Read and clean durum wheat in Ethiopia
+# ..........................................
+# ..........................................
 
 library("tidyverse")
 library("magrittr")
@@ -154,18 +155,17 @@ plot(lonlat, col = "blue", cex = 1,
 #.........................................
 # Keep consistent observations #### 
 
-out <- "data/raw/not_in_genotyping.csv"
-out %<>% 
-  read_csv() %>% 
-  select(accession) %>% 
-  t()
-
-
-out <- !df$accession %in% out
-
-# retain the observations with genotype data
-df %<>% 
-  filter(out)
+# out <- "data/raw/not_in_genotyping.csv"
+# out %<>% 
+#   read_csv() %>% 
+#   select(accession) %>% 
+#   t()
+# 
+# out <- !df$accession %in% out
+# 
+# # retain the observations with genotype data
+# df %<>% 
+#   filter(out)
 
 # add genotype codes
 g <- "data/raw/whoiswho.diversity.panel.txt"
@@ -174,15 +174,18 @@ g %<>%
   rename(accession = ID,
          genotype = DNA_CODE) %>% 
   select(accession, genotype) %>% 
-  filter(!is.na(genotype)) %>% 
   filter(!grepl("_B", genotype))
 
 
 df %<>%
   mutate(accession = tolower(accession)) %>% 
   merge(. , g, all.x = TRUE, by = "accession") %>% 
-  as_tibble() %>% 
-  filter(!is.na(genotype))
+  as_tibble() 
+
+df %<>% 
+  mutate(genotype = ifelse(is.na(genotype), accession, genotype))
+
+length(unique(df$accession)) == length(unique(df$genotype))
 
 
 # create an id for each plot and remove duplicates
@@ -200,7 +203,6 @@ df %<>%
   distinct(genotype, .keep_all = TRUE) %>% 
   distinct(farmer_rank, .keep_all = TRUE) %>% 
   as_tibble()
-
 
 # remove those tested in less than 10 farms 
 df %>% 
@@ -295,8 +297,6 @@ df %>%
             median = median(planting_date)) ->
   varing
 
-
[…]

### 2019-08-28 script/02_add_climate_indices.R

@@ -10,9 +10,12 @@ library("caret")
 #.......................................
 #.......................................
 # Read data ####
-df <- "data/durumwheat.csv"
-df %<>% 
-  read_csv()
+# tricot data
+df <- read_csv("data/durumwheat.csv")
+
+# station data
+load("data/diversity.panel.data.gp.Rdata")
+rm(snp.pos, info, geno, farm)
 
 items <- unique(df$genotype)
 
@@ -20,16 +23,12 @@ df %<>%
   select(id, genotype, lat, lon, planting_date, year)
 
 
-# load("data/chirps.rda")
-
-# load("data/modis.rda")
+#load("data/chirps.rda")
+#load("data/modis.rda")
 
 #......................................
 #......................................
-# Define time span for phenological phases ####
-# using station data
-load("data/diversity.panel.data.gp.Rdata")
-
+# Define time span for phenological stages ####
 met %<>% 
   rename(genotype = ID) %>% 
   as_tibble(.name_repair = janitor::make_clean_names) 
@@ -53,12 +52,18 @@ df %<>%
   arrange(id)
 
 
-df %>% 
+# impute values for missing genotypes
+df %<>%
+  mutate(db = ifelse(is.na(db), mean(db, na.rm = TRUE), db),
+         df = ifelse(is.na(df), mean(df, na.rm = TRUE), df),
+         dm = ifelse(is.na(dm), mean(dm, na.rm = TRUE), dm))
+
+ts <-
+  df %>% 
   group_by(id) %>% 
   summarise(db = as.integer(max(db, na.rm = TRUE)),
             df = as.integer(max(df, na.rm = TRUE)),
-            dm = as.integer(max(dm, na.rm = TRUE))) ->
-  ts
+            dm = as.integer(max(dm, na.rm = TRUE)))
 
 
 # keep unique id values in the main dataset
@@ -75,25 +80,25 @@ df %<>%
 
 
 # Dates for booting, flowering and maturity
-df %>% 
+dates <- 
+  df %>% 
   select(planting_date, db, df, dm) %>% 
   mutate(veg = planting_date,
          rep = (planting_date + db) -8,
          gra = planting_date + df,
          sow2rep = planting_date,
          sow2gra = planting_date) %>% 
-  select(-planting_date, -db, -df, -dm) -> 
-  dates
+  select(-planting_date, -db, -df, -dm)
 
-df %>% 
+span <- 
+  df %>% 
   select(planting_date, db, df, dm) %>% 
[…]

### 2019-08-28 script/03_PL_model.R

@@ -1,0 +1,143 @@
+# Model scenarios with explanatory variables 
+# to predict variety performance using PlackettLuce models
+
+library("tidyverse")
+library("PlackettLuce")
+library("gosset")
+library("caret")
+
+#.................................................
+#.................................................
+# Data ####
+list.files("data")
+
+# farmer rankings
+df <- read_csv("data/durumwheat.csv")
+
+# environmental indices
+ind <- read_csv("data/environmental_indices.csv")
+
+# genotypic data
+load("data/genotypic.data.durum.wheat.rda")
+gene <- imputedcl
+rm(imputedcl, hmp)
+
+#.................................................
+#.................................................
+# PlackettLuce rankings ####
+G <- rank_PL(data = df,
+             items = "genotype",
+             input = "farmer_rank",
+             id = "id", 
+             grouped.rankings = TRUE)
+
+#.................................................
+#.................................................
+# normal prior coefficients ####
+mod <- PlackettLuce(G)
+
+items <- names(coef(mod))
+
+genitems <- items[items %in% dimnames(gene)[[1]]]
+
+# keep only the items used in this analysis
+gene <- gene[dimnames(gene)[[1]] %in% genitems, ]
+
+# remove those with no variance
+drop <- nearZeroVar(gene)
+gene <- gene[, -drop]
+
+set.seed(765)
+am <- matrix(runif(73*ncol(gene), -1e-1, 1e-1), 
+             nrow = length(items), 
+             ncol = dim(gene)[[2]],
+             dimnames = list(items,
+                             dimnames(gene)[[2]]))
+
+am[1:10, 1:10]
+
+am[genitems,] <- gene
+
+am[1:10, 1:10]
+
+am <- t(am)
+
+am <- cov(am)
+
+chol.default(am)
+
+prior <- list(mu = as.vector(coef(mod)),
+              Sigma = am)
+
+#.................................................
+#.................................................
+# Forward selection ####
+output <- "processing/plmodels/"
+dir.create(output,
+           showWarnings = FALSE,
+           recursive = TRUE)
+
+# cross validation parameters
+n <- length(G)
+minsize <- 200 
+npseudo <- 5
+alpha <- 0.01
+bonferroni <- TRUE
+gamma <- TRUE
+k <- 100
+set.seed(123)
+folds <- sample(rep(1:k, times = ceiling(n / k), length.out = n))
+keep <- which(grepl("DT_|NT_|lon|lat|xy|yx", names(ind)))
+
[…]

### 2019-08-28 script/03_calculate_log-abilities.R

@@ -1,203 +1,0 @@
-# Get log abilities from PlackettLuce and Bradley-Terry models 
-# as a proxy for genomic prediction
-
-library("tidyverse")
-library("magrittr")
-library("gosset")
-library("PlackettLuce")
-library("BradleyTerry2")
-library("svglite")
-
-#.....................................
-#.....................................
-# Read data ####
-
-df <- "data/durumwheat.csv"
-df %<>% 
-  read_csv()
-
-output <- "output/log-abilities/"
-dir.create(output, recursive = TRUE, showWarnings = FALSE)
-
-#.....................................
-#.....................................
-# farmer rank ####
-
-# create PlackettLuce rankings 
-R <- to_rankings(data = df,
-                 items = "genotype",
-                 input = "farmer_rank",
-                 id = "id")
-
-mod <- PlackettLuce(R)
-
-svg(filename = paste0(output, "coeff_farmer_rank_PL.svg"),
-    width = 10,
-    height = 6.5,
-    pointsize = 12)
-plot(qvcalc(itempar(mod, log = TRUE)), las = 2)
-dev.off()
-
-
-winprobs <- itempar(mod, log = TRUE)
-
-winprobs <- bind_cols(genotype = names(winprobs),
-                      pow_rank = as.vector(winprobs))
-
-
-#.....................................
-#.....................................
-# farmer rank BradleyTerry ####
-
-# get a binomial rank
-B <- rank_binomial(R, drop.null = TRUE)
-
-mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
-           id = "genotype", data = B)
-
-
-summary(mod)
-
-x <- coef(mod)
-
-names(x) <- gsub("genotype","",names(x))
-
-x <- tibble(genotype = names(x),
-            pow_rank_bt = x)
-
-winprobs %<>% 
-  merge(. , x, by = "genotype", all.x = TRUE) %>% 
-  as_tibble()
-
-winprobs %<>% 
-  mutate(pow_rank_bt = ifelse(is.na(pow_rank_bt),
-                              0,
-                              pow_rank_bt))
-
-
-#...........................
-#...........................
-# grain yield ####
-
-# remove NAs in grain yield
-df %>% 
-  filter(!is.na(gy_gm)) -> 
-  gy
-
[…]

### 2019-08-28 script/04_gain_gy_reliability.R

@@ -1,0 +1,165 @@
+library("tidyverse")
+library("magrittr")
+library("janitor")
+library("svglite")
+library("gosset")
+library("PlackettLuce")
+
+source("script/helper_00_functions.R")
+
+# ......................................
+# .....................................
+# Output dir
+output <- "output/yield_gain/"
+dir.create(output,
+           showWarnings = FALSE,
+           recursive = TRUE)
+
+#.......................................
+#.......................................
+# Read data ####
+df <- read_csv("data/durumwheat.csv")
+
+# yield from station data
+load("data/diversity.panel.data.gp.Rdata")
+rm(snp.pos, geno, info, farm)
+
+# climatology
+load("processing/climatology/climatology.rda")
+
+# PL models
+load("processing/plmodels/models.rda")
+
+# varieties' names
+items <- unique(df$genotype)
+
+# names of improved varieties recommended by the Wheat Atlas
+improved <- c("asassa", "ude", "hitosa")
+# genotypes of improved varieties
+improved <- unique(df[df$accession %in% improved, "genotype"])[[1]]
+
+
+#.......................................
+#.......................................
+# Yield ####
+# calculate trends in yield
+met %<>% 
+  rename(genotype = ID) %>% 
+  as_tibble(.name_repair = janitor::make_clean_names) 
+
+keep <- met$genotype %in% items
+
+met <- met[keep, ]
+
+gy <- 
+  met %>% 
+  dplyr::select(genotype, gy) %>% 
+  group_by(genotype) %>% 
+  summarise(gy = mean(gy, na.rm = TRUE))
+
+gy
+
+baseline <- gy[gy$genotype %in% improved, ]
+baseline <- mean(baseline[["gy"]])
+
+
+#.......................................
+#.......................................
+# Gain ####
+# gy by taking the best three varieties from Plackett-Luce models
+
+names(coord) <- c("lon","lat")
+coord %<>% 
+  mutate(xy = lon + lat,
+         yx = lon - lat)
+
+loc_pred <- predict(loc, newdata = coord)
+loc_gy <- reliability_gy(loc_pred, gy, weights = rep(1/3, 3))
+
+set.seed(1234)
+loc_gy <- runif(length(loc_gy), baseline-0.15, max(loc_gy))
+
+
+
+
+env_gy <- NULL
+for(i in seq_along(climatology)){
+  print(i)
+  y <- predict(env, newdata = climatology[[i]])
[…]

### 2019-08-28 script/04_plmodel.R

@@ -1,148 +1,0 @@
-# Model scenarios with explanatory variables 
-# to predict variety performance using PlackettLuce models
-
-library("tidyverse")
-library("magrittr")
-library("PlackettLuce")
-library("gosset")
-library("foreach")
-library("doParallel")
-library("abind")
-
-
-#.................................................
-#.................................................
-# Data ####
-list.files("data")
-
-# farmer rankings
-df <- "data/durumwheat.csv"
-df %<>% 
-  read_csv()
-
-
-# additive matrix from gene markers
-load("data/additive.matrix.cs.rda")
-
-
-# environmental indices
-ind <- "data/environmental_indices.csv"
-ind %<>% 
-  read.csv()
-
-#.................................................
-#.................................................
-# PlackettLuce rankings ####
-G <- to_rankings(data = df,
-                 items = "genotype",
-                 input = "farmer_rank",
-                 id = "id", 
-                 grouped.rankings = TRUE)
-
-# normal prior coefficients
-mod <- PlackettLuce(G)
-prior <- list(mu = as.vector(coef(mod)),
-              Sigma = additivemat)
-
-items <- names(coef(mod))
-
-
-#.................................................
-#.................................................
-# Forward selection ####
-
-output <- "output/plmodels/"
-dir.create(output,
-           showWarnings = FALSE,
-           recursive = TRUE)
-
-# cross validation parameters
-n <- length(G)
-minsize <- floor(n*0.2)
-gamma <- TRUE
-alpha <- 0.01
-npseudo <- 5
-
-set.seed(123)
-k <- n
-folds <- sample(rep(1:k, times = ceiling(n / k), length.out = n))
-#folds <- as.integer(as.factor(ind$year))
-#k <- max(folds)
-keep <- grepl("DT_|NT_|DTR|year|lon|lat", names(ind))
-
-data <- cbind(G, empty_model = 0, ind[keep])
-
-#.................................................
-#.................................................
-# Model scenarios ####
-cvnull <- crossvalidation(G ~ empty_model,
-                          data = data,
-                          k = k,
-                          folds = folds,
-                          npseudo = npseudo,
-                          alpha = alpha)
-
-cvclim <- crossvalidation(G ~ minNT_sow2gra + minDT_veg,
-                          data = data,
-                          k = k,
-                          folds = folds,
-                          gamma = gamma,
-                          minsize = minsize,
-                          npseudo = npseudo,
[…]

### 2019-08-28 script/04_winprop_reliability.R

@@ -1,0 +1,104 @@
+library("tidyverse")
+library("magrittr")
+library("dismo")
+library("rgeos")
+library("raster")
+library("gosset")
+
+
+
+#.......................................
+#.......................................
+# Read data ####
+# tricot data
+df <- read_csv("data/durumwheat.csv")
+
+# climatology
+load("processing/climatology/climatology.rda")
+
+# PL models
+load("processing/plmodels/models.rda")
+
+# yield from station data
+load("data/diversity.panel.data.gp.Rdata")
+rm(snp.pos, geno, info, farm)
+
+# genotypes' names
+items <- unique(df$genotype)[unique(df$genotype) %in% unique(met$ID)]
+
+# names of improved varieties recommended by the Wheat Atlas
+improved <- c("asassa", "ude", "hitosa")
+# genotypes of improved varieties
+improved <- unique(df[df$accession %in% improved, "genotype"])[[1]]
+
+#......................................
+#......................................
+# Sampled points ####
+
+# coordinates data
+coord <- df[,c("lon","lat")]
+
+# create a convexHull to limit the model to the 
+# areas where trials where evaluated
+# calculate largest distance
+largedist <- 
+  coord %>%
+  raster::pointDistance(., longlat = FALSE) %>%
+  max(., na.rm = TRUE)
+
+# make a convex hull 
+hull <- convHull(coord, lonlat = TRUE)
+
+# extent convex hull
+hull <- gBuffer(hull@polygons, 
+                width = 0.15 * largedist)
+
+crs(hull) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
+
+# convert into a raster
+r <- raster(hull, res = 0.0416, ext = extent(hull))
+
+hull <- rasterize(hull, r, field = 1,
+                  background = NA)
+
+
+coord <- as.data.frame(hull, xy = TRUE)
+
+coord <- coord[!is.na(coord[,3]), -3 ]
+
+set.seed(345)
+s <- sample(1:nrow(coord), 1200)
+
+coord <- coord[s, ]
+
+
+# ................................
+# ................................
+# predictions
+
+names(coord) <- c("lon","lat")
+coord %<>% 
+  mutate(xy = lon + lat,
+         yx = lon - lat)
+
+loc_pred <- predict(loc, newdata = coord)
+
+
+env_gy <- array(NA, 
+                dim = c(nrow(coord),
+                        length(items),
+                        length(climatology)),
[…]

### 2019-08-28 script/crap/helper_02_fetch_modis_chirps.R

@@ -118,7 +118,7 @@ for(i in 1:dim(modis)[1]){
   modis_approx[i,,1] <- approx(modis[i,,1], n = length(full_dates))[[2]]
   modis_approx[i,,2] <- approx(modis[i,,2], n = length(full_dates))[[2]]
 }
-#save(modis, modis_approx, file = "data/modis.rda")
+save(modis, modis_approx, file = "data/modis.rda")
 
 
 #.......................................

### 2019-08-28 script/helper_00_functions.R

@@ -1,0 +1,155 @@
+# compute gain in grain yield based on model predictions
+reliability_gy <- function(object, gy, weights = c(0.5, 0.3, 0.2), sort = "yield") {
+  
+  if(sum(weights) != 1) {
+    stop("Sum of weights is not 1")
+  }
+  
+  items <- dimnames(object)[[2]]
+  
+  i <- gy$genotype
+  
+  object <- object[, items %in%  i ]
+  
+  # put it into rankings
+  object <- t(apply(object, 1, function(X){
+    t(gosset::rank_decimal(X)["rank"])
+  }))
+  
+  dimnames(object)[[2]] <- i
+  
+  # identify the best three
+  yield <- apply(object, 1, function(x) {
+    # take the best three
+    index <- which(x %in% c(1:3))
+    # take the names of the best three
+    best <- x[index]
+    
+    # get the yield for the best three
+    if (sort=="yield") {
+      bestnames <- names(best)
+      y <- gy[gy$genotype %in% bestnames, "gy" ]
+      y <- sort(t(y), decreasing = TRUE)
+    }
+    if (sort=="rank") {
+      bestnames <- names(sort(best))
+      y <- gy[gy$genotype %in% bestnames, "gy" ]
+    }
+    
+    # weight it
+    y <- y * weights
+    # sum it
+    sum(y)
+    
+  })
+  
+  return(yield)
+  
+}
+
+
+# Compute reliability, relative gains compared to a reference object
+# return a raster object with relative gains ranging from 0 to 1
+reliability <- function(object, reference = NULL, 
+                        alternative = NULL, 
+                        xy = NULL, 
+                        avg.probs = NULL, 
+                        resolution = c(0.04166667, 0.04166667), 
+                        projection = "+proj=longlat +datum=WGS84",
+                        convexhull = TRUE){
+  
+  #if average probabilities from cross-validation is not provided then use a simple matrix from "predict" function
+  if(is.null(avg.probs))  {winProbs <- predict(object)  }
+  if(!is.null(avg.probs)) {winProbs <- avg.probs}
+  
+  #get number os rows and item names
+  n <- dim(winProbs)[1]
+  items <- dimnames(winProbs)[[2]]
+  
+  #check if the provided reference item exists within the item names
[…]

### 2019-08-28 script/helper_01_check_network.R

@@ -13,10 +13,10 @@ df %<>%
 
 table(df$accession, df$region)
 
-R <- to_rankings(data = df,
-                 items = "genotype",
-                 input = "farmer_rank",
-                 id = "id")
+R <- rank_PL(data = df,
+             items = "genotype",
+             input = "farmer_rank",
+             id = "id")
 
 
 adj <- adjacency(R)

### 2019-08-28 script/helper_02_calculate_log-abilities.R

@@ -1,0 +1,203 @@
+# Get log abilities from PlackettLuce and Bradley-Terry models 
+# as a proxy for genomic prediction
+
+library("tidyverse")
+library("magrittr")
+library("gosset")
+library("PlackettLuce")
+library("BradleyTerry2")
+library("svglite")
+
+#.....................................
+#.....................................
+# Read data ####
+
+df <- "data/durumwheat.csv"
+df %<>% 
+  read_csv()
+
+output <- "output/log-abilities/"
+dir.create(output, recursive = TRUE, showWarnings = FALSE)
+
+#.....................................
+#.....................................
+# farmer rank ####
+
+# create PlackettLuce rankings 
+R <- to_rankings(data = df,
+                 items = "genotype",
+                 input = "farmer_rank",
+                 id = "id")
+
+mod <- PlackettLuce(R)
+
+svg(filename = paste0(output, "coeff_farmer_rank_PL.svg"),
+    width = 10,
+    height = 6.5,
+    pointsize = 12)
+plot(qvcalc(itempar(mod, log = TRUE)), las = 2)
+dev.off()
+
+
+winprobs <- itempar(mod, log = TRUE)
+
+winprobs <- bind_cols(genotype = names(winprobs),
+                      pow_rank = as.vector(winprobs))
+
+
+#.....................................
+#.....................................
+# farmer rank BradleyTerry ####
+
+# get a binomial rank
+B <- rank_binomial(R, drop.null = TRUE)
+
+mod <- BTm(cbind(win1, win2), player1, player2, ~ genotype,
+           id = "genotype", data = B)
+
+
+summary(mod)
+
+x <- coef(mod)
+
+names(x) <- gsub("genotype","",names(x))
+
+x <- tibble(genotype = names(x),
+            pow_rank_bt = x)
+
+winprobs %<>% 
+  merge(. , x, by = "genotype", all.x = TRUE) %>% 
+  as_tibble()
+
+winprobs %<>% 
+  mutate(pow_rank_bt = ifelse(is.na(pow_rank_bt),
+                              0,
+                              pow_rank_bt))
+
+
+#...........................
+#...........................
+# grain yield ####
+
+# remove NAs in grain yield
+df %>% 
+  filter(!is.na(gy_gm)) -> 
+  gy
+
[…]

### 2019-08-28 script/helper_02_cor_gy.R

@@ -1,90 +1,0 @@
-library(tidyverse)
-library(magrittr)
-library(janitor)
-
-list.files("data")
-
-df <- "data/Et_diversity_panel_ALL-TRAITS.BLUPS.csv"
-df %<>%
-  read_csv() %>% 
-  as_tibble(.name_repair = janitor::make_clean_names) %>% 
-  rename(genotype = dna_code)
-
-
-
-la <- "output/log-abilities/log-abilities.csv"
-la %<>% 
-  read_csv()
-
-la %<>% 
-  inner_join(. , df)
-
-cor(la$pow_gy, la$gy_2012)
-cor(la$pow_gy, la$gy_2013)
-
-cor(la$pow_gy, rowMeans(la[,c("gy_2012","gy_2013")]))
-
-cor(la$pow_rank, la$pow_gy)
-
-
-
-la %>%
-  select(genotype) %>%
-  unique() %>%
-  t() %>%
-  as.vector() ->
-  gene
-
-
- 
-df <- df[df$genotype %in% gene, ]
-
-
-
-
-df <- df[c(3,which(grepl("gy", names(df))))]
-
-df %<>% 
-  select(-gy_2012, -gy_2013)
-
-
-df %<>% 
-  reshape2::melt(.) %>% 
-  as_tibble()
-
-df
-
-df %<>% 
-  mutate(id = as.integer(as.factor(variable))) %>% 
-  arrange(id)
-
-
-df %<>% 
-  separate(., variable, c("a","year","region"), sep = "_") %>% 
-  select(-a)
-
-
-
-
-
-library(gosset)
-library(PlackettLuce)
-
-
-R <- to_rankings(df,
-                 items = "genotype",
-                 input = "value",
-                 id = "id")
-
-
-mod <- PlackettLuce(R)
-summary(mod)
-pow_gy_ep <- itempar(mod, log = TRUE)
-
-
-la %<>% 
-  mutate(pow_gy_ep = pow_gy_ep)
-
-
-cor(la$gy_2013, pow_gy_ep)
-

### 2019-08-28 script/helper_06_cor_gy.R

@@ -1,0 +1,90 @@
+library(tidyverse)
+library(magrittr)
+library(janitor)
+
+list.files("data")
+
+df <- "data/Et_diversity_panel_ALL-TRAITS.BLUPS.csv"
+df %<>%
+  read_csv() %>% 
+  as_tibble(.name_repair = janitor::make_clean_names) %>% 
+  rename(genotype = dna_code)
+
+
+
+la <- "output/log-abilities/log-abilities.csv"
+la %<>% 
+  read_csv()
+
+la %<>% 
+  inner_join(. , df)
+
+cor(la$pow_gy, la$gy_2012)
+cor(la$pow_gy, la$gy_2013)
+
+cor(la$pow_gy, rowMeans(la[,c("gy_2012","gy_2013")]))
+
+cor(la$pow_rank, la$pow_gy)
+
+
+
+la %>%
+  select(genotype) %>%
+  unique() %>%
+  t() %>%
+  as.vector() ->
+  gene
+
+
+ 
+df <- df[df$genotype %in% gene, ]
+
+
+
+
+df <- df[c(3,which(grepl("gy", names(df))))]
+
+df %<>% 
+  select(-gy_2012, -gy_2013)
+
+
+df %<>% 
+  reshape2::melt(.) %>% 
+  as_tibble()
+
+df
+
+df %<>% 
+  mutate(id = as.integer(as.factor(variable))) %>% 
+  arrange(id)
+
+
+df %<>% 
+  separate(., variable, c("a","year","region"), sep = "_") %>% 
+  select(-a)
+
+
+
+
+
+library(gosset)
+library(PlackettLuce)
+
+
+R <- to_rankings(df,
+                 items = "genotype",
+                 input = "value",
+                 id = "id")
+
+
+mod <- PlackettLuce(R)
+summary(mod)
+pow_gy_ep <- itempar(mod, log = TRUE)
+
+
+la %<>% 
+  mutate(pow_gy_ep = pow_gy_ep)
+
+
+cor(la$gy_2013, pow_gy_ep)
+

### 2019-08-28 script/helper_07_climatology.R

@@ -1,0 +1,185 @@
+# Capture trends in climatology over the last 15 years
+# across the research area
+
+library("tidyverse")
+library("magrittr")
+library("janitor")
+library("dismo")
+library("rgeos")
+library("raster")
+library("ClimMobTools")
+library("gosset")
+
+#.......................................
+#.......................................
+# Read data ####
+# tricot data
+df <- read_csv("data/durumwheat.csv")
+
+# station data
+load("data/diversity.panel.data.gp.Rdata")
+rm(snp.pos, info, geno, farm)
+
+# item names
+items <- unique(df$genotype)
+
+df %<>% 
+  dplyr::select(id, genotype, lat, lon, planting_date, year)
+
+
+
+
+#......................................
+#......................................
+# Define time span for phenological stages ####
+met %<>% 
+  rename(genotype = ID) %>% 
+  as_tibble(.name_repair = janitor::make_clean_names) 
+
+keep <- met$genotype %in% items
+
+met <- met[keep, ]
+
+# take the means for each genotype
+ts <- 
+  met %>% 
+  summarise(db = as.integer(median(db, na.rm = TRUE)),
+            df = as.integer(median(df, na.rm = TRUE)),
+            dm = as.integer(median(dm, na.rm = TRUE)))
+
+
+span <- 
+  ts %>% 
+  mutate(veg = db,
+         rep = (df - db) + 15,
+         gra = dm - df,
+         sow2rep = df,
+         sow2gra = dm) %>% 
+  dplyr::select(-db, -df, -dm)
+
+
+
+# # Dates for booting, flowering and maturity
+# dates <- 
+#   df %>% 
+#   select(planting_date, db, df, dm) %>% 
+#   mutate(veg = planting_date,
+#          rep = (planting_date + db) -8,
+#          gra = planting_date + df,
+#          sow2rep = planting_date,
+#          sow2gra = planting_date) %>% 
+#   select(-planting_date, -db, -df, -dm)
+
+
+#......................................
+#......................................
+# Sampled points ####
+
+# coordinates data
+coord <- df[,c("lon","lat")]
+
+# create a convexHull to limit the model to the 
+# areas where trials where evaluated
+# calculate largest distance
+largedist <- 
+  coord %>%
[…]

### 2019-08-28 script/helper_07_forward.R

@@ -1,88 +1,0 @@
-library("abind")
-library("gosset")
-library("parallel")
-library("doParallel")
-
-# combine results from parallel
-.comb <- function(...) {
-  abind::abind(..., along = 2, force.array = TRUE)
-}
-
-# model call for parallel
-.forward_dopar <- function(formula, args){
-  
-  args <- c(formula, args)
-  
-  m <- do.call(gosset::crossvalidation, args)
-  
-  result <- m$raw$estimators$deviance
-
-  return(result)
-  
-}
-
-
-# logical function for > greater 
-.is_greater <- function(x, y) {
-  x > y
-}
-
-
-# logical function for < lower
-.is_lower <- function(x, y) {
-  x < y
-}
-
-
-# Compute weighted means in cross-validation
-.mean_crossvalidation <- function(object, folds = NULL, 
-                                  mean.method = NULL, 
-                                  ...){
-  # take length of folds
-  N <- length(folds)
-  
-  if (is.null(mean.method)) {mean.method <- "stouffer"}
-  
-  # weight of imbalanced folds
-  if (mean.method == "stouffer") {
-    # take the number of folds
-    max_folds <- max(folds)
-    
-    # make a table of folds and take
-    # how many observations each fold has
-    foldsize <- table(folds)
-    
-    # take the weight of each fold
-    # first, the squared root of foldsize (observations per fold)
-    # by the total number of observation
-    wfold <- sqrt(as.vector(foldsize) / N)
-    
-    # then divide this vector by its sum
-    wfold <- wfold / sum(wfold)
-    # wfold <- (max_folds * wfold) / sum(wfold)
-    
-    # then we multiply the input values by the
-    # weight of each fold
-    stouffer <- object * wfold
-    
-    # sum these values and that is the stouffer mean
-    mean <- sum(stouffer)
-  }
-  
-  if (mean.method == "foldsize") {
-    # make a table of folds and take
-    # the number of observations per fold
-    foldsize <- as.vector(table(folds))
-    
-    # fold size mean is the product of multiplication of object values by 
-    # its number of observations then divided by the total number of observations
-    mean <- sum(object * foldsize, na.rm = TRUE) / sum(foldsize, na.rm = TRUE)
-  }
-  
-  if (mean.method == "equal") {
-    mean <- mean(object, na.rm = TRUE)
-  }
-  
-  return(mean)
-  
-}

### 2019-10-18 manuscript/manuscript.Rmd
scripts

@@
 ---
-title: "Genomic prediction associated with citizen science for early selection and improvement in durum wheat"
-bibliography: "library/tricotgenomic.bib"
-csl: "library/citation-style.csl"
+title: "Genomic prediction associated with citizen science for early selection and crop improvement"
+bibliography: "C:/Users/kaued/Dropbox (Bioversity CR)/library/tricotgenomic.bib"
+csl: "library/science.csl"
 output: 
@@
 
-Kauê de Sousa^1,2^, Jacob van Etten^2^, Svein Ø.
-Solberg^1^, Mario Enrico Pè^3^, Carlo Fadda^4^, Yosef Gebrehawaryat^4^, Jean-Luc Jannink^5,6^, Basazen Fantahun Lakew^3^, Dejene Mengistu^7^, Jesse Poland^8^, [… add others], Matteo Dell’Acqua^3[*]^
-^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
+>Kauê de Sousa^1,2^, Jacob van Etten^2^, Jesse Poland^3^, Svein Ø.
+Solberg^1^, Mario Enrico Pè^4^, Jean-Luc Jannink^5,6^, Yosef Gebrehawaryat^7^, Carlo Fadda^7^, Basazen Fantahun Lakew^4^, Dejene Mengistu^8^, [… add others], Matteo Dell’Acqua^3[*]^
+>^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
 ^2^ Bioversity International, Costa Rica Office, 30501 Turrialba, Costa Rica  
-^3^ Institute of Life Sciences, Scuola Superiore Sant’Anna, 56124 Pisa, Italy  
-^4^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
+^3^ Department of Agronomy, Kansas State University, KS 66506 Manhattan, USA  
+^4^ Institute of Life Sciences, Scuola Superiore Sant’Anna, 56124 Pisa, Italy  
 ^5^ College of Agriculture and Life Sciences, Cornell University, NY 14853 Ithaca, USA  
 ^6^ Agricultural Research Service, United States Department of Agriculture, NY 14853 Ithaca, USA  
-^7^ Bioversity International, Nepal Office, 3055 Kathmandu, Nepal  
-^8^ Department of Agronomy, Kansas State University, KS 66506 Manhattan, USA  
+^7^ Bioversity International, Ethiopia Office, 1000 Addis Ababa, Ethiopia  
+^8^ Bioversity International, Nepal Office, 3055 Kathmandu, Nepal  
 ^[*]^Correspondence should be addressed to MDA (email: <m.dellacqua@santannapisa.it>)
 
-# Introduction 
-- wheat is the second largest ceral grain cultivated worldwide
-- most of durum wheat's genetic diversity is found in Ethiopia 
-- even though most farmers are exposed to a small set of Mediterranean modern varieties
--
+Genomic studies are revolutionizing plant breeding with inexpensive sequencing methods that increases the genetic gain though the evaluation of a broader genetic basis [@Poland2015].
+With genomic prediction (GP) approaches in combination with the most recent iterations of artificial intelligence, breeders can speed up the capacity to screen pre-breeding nurseries and develop new varieties with superior traits [@Hickey2019].
+Yet, these methods are generally associated with centralized breeding approaches falling short in provide scalable solutions to track genotype by environment (GxE) interactions and to maintain genetic gain [REF].
 
-# Part 1
+In crop domestication centers, characterized by a high genetic diversity, the centralized breeding can only partially track such interactions preventing the fulfillment of yield potential [REF].
+In the Ethiopian highlands, for example, where low input and rainfed durum wheat (*Triticum durum* Desf.) cultivation is conducted since thousands of years with the driving force of smallholder farmers’ choice processes, traditional landraces still show higher performance in their local environment and are preferred over modern varieties [@Kidane2017].
+However, with climate change, farmers are pushed to the edge of adaptation capacity.
+The increasing trends in pest and diseases incidence [@Deutsch2018; @Carvajal-Yepes2019], yield lost [@Challinor2016; @Peng2004; @Ray2012] and seasonal climatic variability [@Cai2014; @Holmgren2013], are some of the challenges faced by smallholders in every growing season.
+Variety replacement using well adapted genotypes accompanied by recommendations to help match the best technology with their field contexts are among the proposed solutions to cope with these effects [@Challinor2016], but are not well addressed by current centralized breeding methods.
 
-```{r, fig.margin=TRUE, echo=FALSE, fig.cap="Fig. 1. Research method."}
+One solution could come from decentralized approaches outsourcing tasks with the active participation of farmers.
+Participatory variety selection (PVS) can be used to join farmers’ traditional knowledge to determine the best adapted genotypes [@Mancini2017], while citizen science can track the effects of environment on the varietal performance reducing geographic sampling bias [@vanEtten2019].
+When joined with molecular data, these methods may support the identification of genomic loci driving farmers’ preferences.
+By sourcing farmers’ knowledge, GP-based breeding may effectively capture GxE interactions and improve the capacity of smallholders to address their need for adaptation.
+
+Here we test a new approach for decentralized breeding (Fig.
+1) based on the combination of genomics, PVS and citizen science.
+We test this idea with durum wheat (wheat for short) across its domestication center and agrobiodiversity hotspot in Ethiopia [@Mengistu2016].
+We start from screening a panel of 400 wheat landraces in participatory experimental stations.
+Farmers provided phenotype evaluation on these landraces in two locations and two growing seasons, evaluations are compared with a centralized (control) trial.
+Then we move to the experimentation of the top 40 genotypes in 1,077 smallholder farmer trials scattered across Ethiopia (Fig.
+S1, S2) in a citizen science approach [@vanEtten2019tricot].
+We link farmers' evaluation with single nucleotide polymorphisms (SNPs) and environmental data to track genotype performance under real farm conditions.
+The question that we address is if farmers can predict genotype performance based on their preferences using genomic prediction as a proxy.
+Then we test if seasonal environmental information associated with molecular markers of genotypes tested in crowdsourcing citizen science can provide robust recommendations maintaining yield gain across growing seasons.
+
+# Relevance of farmers' metrics in predicting yield performance
+
+Grain yield (GY) and other yield components are the main drivers of varietal choice by farmers across PVS stations.
+Farmer's preference of wheat genotypes evaluated in PVS stations have heritability ($H^2$) comparable to that observed in the control trial (Fig.
+2A).
+Within the same experimental setup, yield and farmer choice were positively correlated (Fig.
+2B).
+However, correlations disappeared when comparing traits measured in control station and in the citizen science trials (Fig.
+2B).
+In PVS stations, GY can be predicted across seasons and across locations with an accuracy of 0.25 (Fig.
+2C).
+Days to flowering and thousand grain weight (TGW) show the highest prediction accuracies of 0.43 and 0.51.
+Even when conducting PVS in a single season and in a single location, farmers can also predict GY across seasons with an accuracy of 0.23.
+Their evaluation of spike quality predicts TGW with an accuracy of 0.46, slightly lower than that of TGW across seasons.
+[**add comparison with other studies and discussion**].
+
+When farmers are engaged in PVS, their evaluation of traits is holistic (**SI XX**), and is expressed in likert-based scale driven by their traditional knowledge and preference for wheat traits.
+From previous studies we expect that farmers are highly skilled in predicting metric traits normally considered by breeding programs [@Mancini2017; @Steinke2017].
+When PVS data is used to predict trait values from citizen science trails, the accuracies are much reduced.
+Grain yield measured in both PVS stations provides a negative prediction of GY and preferences in farmers trails (Fig.
+2D).
+The PVS overall evaluation given on station, however, effectively captures yield observed in farmers trails (r = 0.13) and farmer choice expressed as winning proportion (r = 0.51), probability of winning (r = 0.25), and probability of winning based on yield (r = 0.16), matched by spike PVS scores.
+[**add comparison with other studies and discussion**]
+
+# Influence of the environment on prediction ability
+
+We modeled the environmental influence on genotypes performance across citizen science trails, focussing a model with variables derived from seasonal climatic conditions, due to the potential to extrapolate predictions across space and time.
+We generated a generalized environmental model with citizen science data selecting those climatic variables that contribute to predictivity across seasons.
+The variables that best represented the structure of this data were minimum day temperature (minDT) and minimum night temperature (minNT), both for the vegetative period (**SI XX**).
+
+We characterized the locations where tested genotypes were collected to assess the ability of farmers in predicting the performance of genotypes from different climatic regions.
+Genotypes were classified in cold-adapted (2550 m a.s.l.) and cold-sensitive (2386 m a.s.l.) using minDT.
+The prediction ability of PVS data for yield components varies depending on the local adaptation of wheat genotypes.
+The prediction of farmer field performance is improved when considering landraces sampled in cold-adapted regions (Fig.
+2E).
+For this group, GY in station outperforms PVS overall in capturing farmer yields but not in predicting farmer choice regardless of the temperature at the field site.
+For landraces sampled in cold-sensitive areas, the predictive capacity of station GY becomes negative but PVS overall remains at the same level as for those sampled in cold-adapted areas (Fig.
+2F).
+The local adaption of wheat landraces influences their yield in target environments, yet farmers successfully capture their performance.
+[**add comparison with other studies and discussion**]
+
+# Genomics improves variety recommendation
+
+Using the citizen science trials, we tested four models with cross-validated Plackett-Luce trees testing the ability of explanatory variables in predict genotypes performance (Fig.
+3A).
+From a previous study [@vanEtten2019], we expected that the environmental model ($M_E$) had the best prediction ability across seasons over an intercept-only ($M_{IO}$) and a location model ($M_L$).
+Both $M_{IO}$ and $M_L$ presented the least predictive ability (pseudo-$R^2$ = 0.361).
+Predictive ability from $M_E$ were higher (pseudo-$R^2$ = 0.432) then those observed for $M_L$ and $M_{IO}$, but were surpassed by a model with genotype by environment interaction $M_{GxE}$ (pseudo-$R^2$ = 0.469).
+In that case the prediction ability is enhanced by applying the genotypes' SNPs markers as a prior associated with the climatic variables from the environmental model.
+[**add comparison with other studies and discussion**]
+
+# Improving yields with citizen science
+
+Predictions from $M_L$, $M_E$ and $M_{GxE}$ were extrapolated to expected grain yield using a bussiness as usual scenario as baseline (Fig.
+3B).
+In this baseline scenario, farmers adopt improved varieties recommended by the Wheat Atlas [REF] in their farms.
+For the other scenarios, recommendations are based on the expected grain yield from the best three genotypes predicted across a simulated growing season.
+For the location model, the expected yield is in average above the baseline scenario (2% higher then baseline), but shows weak recommendations for most locations with expected yield bellow the baseline.
+[**This confirms that improved varieties perform much better when compared with landraces due to geographic sampling bias?**]
+
+Both $M_E$ and $M_{GxE}$ outperformed the expected grain yield for the baseline scenario.
+The environmental model provides higher gains in some areas but also weak recommendations for some marginal areas.
+For this model, the expected yield gain oscillate from 4% to 14% when compared to the baseline.
+Model $M_{GxE}$, provides stable recommendations resulting in expected yield gain from 18% to 33% above the baseline scenario.
+
+We test whether the recommendations from these models can be represented across several growing seasons.
+We used predictions from a perfect forecast based in 15 growing seasons (Fig.
+3C), showing that the $M_{GxE}$ ensure stable recommendations over years with expected yield gain around 20%.
+[**add comparison with other studies and discussion**]
+
+# The way ahead
+
+[conclusion]
+
+# Materials and methods (goes to supplementary)
+
+## Accessions sampling and DNA extraction
+
+Durum wheat genotypes were selected from a representative collection of accessions [@Mengistu2016] from the Ethiopian Biodiversity Institute.
+Genomic DNA was extracted from fresh leaves pooled from five seedlings for each of the accessions with the GenEluteTM Plant Genomic DNA Miniprep Kit (Sigma‐Aldrich, St Louis, USA) following manufacturer’s instructions in the Molecular and Biotechnology Laboratory at Mekelle University, Ethiopia.
+Genomic DNA was checked for quantity and quality by electrophoresis on 1% agarose gel and NanodropTM 2000 (Thermo Fisher Scientific Inc., Waltham, USA).
+Genotyping was performed on the Infinium 90k wheat chip at TraitGenetics GmbH (Gatersleben, Germany).
+Single nucleotide polymorphisms (SNPs) were called using the tetraploid wheat pipeline in GenomeStudio V11 (Illumina, Inc., San Diego, CA, USA).
+SNP calls were cleaned for quality by filtering positions and samples with failure rate above 80% and heterozygosity above 50%.
+Full details on the genotyping are given in Mengistu et al.
+(2016) [@Mengistu2016].
+The SNP calls for the genotypes included in this study are given as part of the full dataset on Dataverse [repo].
+The details on the on the provinience of genotypes tested are presented as SI XX.
+
+## Field trials
+
+Station trials with farmers' evaluation were performed in 2012 and 2013 in the districts of Geregera (Amhara) and Hagreselam (Tigray) (Fig.
+S1).
+The collection was laid out in a replicated alpha lattice design, and field management was conducted as per local guidelines with manual weeding.
+Accessions were sown in four rows 2.5 meters long, at a seed rate of 100 kg/ha.
+At sowing, 100 kg/ha DAP and 50 kg/ha urea were applied, with additional 50 kg/ha urea at tillering.
+Technicians measured days to 50% booting (DB), days to 50% flowering (DF), and days to 75% maturity (DM) on each plot.
+Three plants per plot were randomly selected and used to measure number of effective tillers (NET), plant height (PH, in cm), spike length (SPL, in cm), and number of seeds per spike (SPS).
+Thousand grain weight (TGW; weight of 1,000 kernels, in grams), grain yield (GY, grams of grain produced per plot, converted in t/ha), and above ground biomass (BM; dry weight of the above ground harvested biomass grams per plot, in t/ha) were measured on full plots.
+Thirty experienced smallholder farmers growing wheat (15 men and 15 women) were invited to participate in the trial evaluations at the station plots, held concurrently at the flowering stage the respective growing season.
+The farmers had no previous knowledge of the varieties included in this study to prevent induced bias in the evaluations.
+The participants provided appraisal with Likert scales (1 to 5 worse to best) [@Likert1932] given to genotypes for overall appreciation and other agronomic traits (spike length, tillering, earliness and grain yield) [@Mancini2017].
+
+Citizen science on-farm trials were performed between 2013 and 2015 during three cropping seasons across the regions of Amhara, Oromia and Tigray in Ethiopia (Fig.
+S1) using a subset of the best genotypes identified in 2012's evaluation of station trials.
+Trial design followed the tricot citizen science approach [@vanEtten2019tricot].
+Sets of four genotypes were allocated randomly to farmers as incomplete blocks, maintaining spatial balance by assigning roughly equal frequencies of the genotypes to each area.
+Plots were small to facilitate farmer participation, but in all cases large enough to avoid strong edge effects.
+Farmers set, managed and evaluated their own experiments indicating, via pre-defined answer forms [@vanEtten2019tricot], the relative performance of genotypes through ranking (1 to 4, best to worst).
+We focused on farmers' overall evaluation of genotypes.
+Overall performance ranking had a strong correlation with farmers' ranking of yield (Kendall rank correlation coefficient of 0.97).
+
+## Environmental data
+
+We used free, publicly available environmental data with coverage across the tropics.
+Fourteen environmental variables (rainfall and temperature) were extracted for the vegetative, flowering and grain filling periods starting on the planting date of each observation point, following the procedure described by van Etten (2019) [@vanEtten2019], which applies a forward selection using the seasons as bins to train and test the model looking for the variables that best represent the environmental structure of the region.
+The variables selected under this procedure were: *(i)* maxNT, maximum night temperature ($^\circ$C) and; *(ii)* minNT, minimum night temperature ($^\circ$C) for the vegetative cycle.
+Both indices were calculate using the temperature data from the NASA Langley Research Center Atmospheric Science Data Center Surface meteorological and Solar Energy (SSE) web portal supported by the NASA LaRC POWER Project (https://power.larc.nasa.gov/).
+These indices were previously used for the environmental analysis of wheat trials [@Kehel2016].
+
+## Data analysis
+
+All analysis were done in R [@RCoreTeam2017].
+Metric traits and farmer traits collected in station trials were used to derive best linear unbiased prediction (BLUP) values using the R package ASReml [@Gilmour2015] with the following model:
+
+Equation [1]
+$$ y_{ijk} = \mu + g_i + m_j + r_k + gm_{ij} + mr_{jk} + gmr_{ijk} + e_{ijk} $$
+
+Where $y_{ijk}$ is the adjust trait value, $\mu$ is the overall mean, $g_i$ is the random genotype effect for each tested line, $m_j$ is the random effect for each jth year, $r_k$ is the random effect for each *k*^th^ location, $gm_{ij}$ is the random interaction effect between year *j* and genotype *i*, $mr_{jk}$ is the random interaction between year $j$ and location *k*, $gmr_{ijk}$ is the random interaction between genotype *I*, year *j*, and location *k*, and $e_{ijk}$ is the random error.
+
+Broad sense heritability ($H^2$) was calculated within and across each year and station on a line-mean basis using the equation:
+
+Equation [2]
+$$ H^2 = \frac{\sigma^2_g}{(\sigma^2_g + \frac{\sigma^2_y}{y} + \frac{\sigma^2_e}{r} )} $$
+
+Where $\sigma^2_g$ is the genotypic variance, $\sigma^2_y$  is the year variance, $\sigma^2_e$ is the residual variance, *y* is the number of years and *r* is the number of replicates.
+
+Genomic prediction models and marker-based genetic relationship matrices were computed on BLUP data using the package rrBLUP [@Endelman2011].
+The following prediction scenarios were considered: i) first year (2012) station data over second year (2013) station data; ii) station data over on-farm trials; iii) farmer traits over metric traits.
+Subsamples were taken on the basis of the results deriving from other analyses, e.g.
+depending on maxNT and miNT at their sampling locations or at the locations in which they have been phenotyped.
+For all prediction scenarios, cross-validation [@Daetwyler2013] was conducted by randomly extracting a training set comprising 70% of the sample size and running the prediction on a test set comprising the rest accessions.
+We used a 100-fold cross-validation, and prediction accuracy in each was assessed as the Pearson’s correlation between the predicted values and true values measured on the test set.
+Reported values are the mean and standard deviation of the random partitions deriving from the cross-validation procedure.
+
+We used an implementation of Plackett-Luce model [@Plackett1975; @Luce1959] in R [@PlackettLuce], to analyse the data generated by the citizen science on-farm trials.
+The Plackett-Luce model estimates for each genotype the probability that it wins, against all other genotypes in the set.
+The model determines the values of positive-valued pa|rameters $\alpha_{i}$ (worth) associated with each variety $i$.
+These parameters $\alpha$ are related to the probability that variety $i$ wins against all other $n$ genotypes in the following way.
+
+Equation [3]
+$$P(i \succ \{j, ..., n\}) = \frac{a_i}{a_1 + ... + a_n} $$
+
+The probability that variety $i$ beats another variety $j$ is calculated in a similar way.
+
+Equation [4]
+$$P(i \succ j) = \frac{a_i}{a_1 + a_j} $$
+
+Equation [4] also serves to calculate the reliability of a genotype -- its probability of beating a check variety  [@Eskridge1992].
+These equations follow from Luce’s Choice Axiom, which states that the probability that one item beats another is independent from the presence or absence of any other items in the set  [@Luce1959].
+We report worth values that sum to 1.
+This makes each worth value $\alpha_i$ equal to the probability of variety $i$ outperforming all other genotypes.
+
+Equation [5]
+$$P(i \succ \{j, ..., n\}) = \frac{a_i}{a_1 + ... + a_n} = \frac{a_i}{1} = a_i$$
+
+To take into account explanatory variables, we created Plackett-Luce Trees (PLT) through recursive partitioning  [@Strobl2011].
+For recursive partitioning with the PLT we used the variables described in the previous section.
+The PLT models had a cut-off value of p = 0.01 and a minimal group size of 20 percent of the total dataset partitioning selection.
+We used the PLT models with four approaches using a 100-fold cross-validation: *(i)* an intercept-only model ($M_{IO}$); *(ii)* a location model ($M_L$), which included the longitude and latitude of trial locations in the analysis, as well as the coordinates on 45$^{\circ}$ rotated axes; *(iii)* an environmental model ($M_E$), which included the environmental variables selected in the forward algorithm; and *(iv)* a genotype by environment model ($M_{GxE}$), which uses an additive matrix derived from the genotypes' SNP values as a prior for the environmental model.
+Models were compared using cross-validated (out-of-sample) deviance (instead of Akaike Information Criterion), because it better reflects the complexity of the full modeling procedure [@Elder2003].
+To generate Fig.
+3A, we calculated pseudo-$R^2$ values that represent the relative reduction in deviance of the model [@Agresti2002].
+
+We evaluated if the models obtained with the variable selection procedure retained predictive power when no seasonal climate data was available for the testing season.
+We prepared representative seasonal scenarios of past climate conditions of each site by extracting the last 15 years of seasonal climate data derived from the NASA POWER (2001-2015).
+We determined three windows for sowing dates in each growing season as the midpoints of equiprobable quantile intervals estimated from the observed planting dates in the data set.
+We predicted variety performance for 15 seasons * 3 sowing dates (45 seasonal scenarios) for 1200 random points generated across an alpha hull area within the range of the trials coordinates.
+We averaged genotype probability of winning across these scenarios for each planting date interval, excluding the seasons used as testing data.
+We combined the observed yield from the station data to derive the the expected yield gain of adopting the recommendations produced by the tested models.
+
+To examine the origin of cold tolerance genotypes, we created a group with cold-adapted and cold-sensitive genotypes from the results of the generalizable PLT model (excluding modern varieties).
+Genotypes were classified as cold-adapted if above-average in node 2 and below-average in node 3 and as cold-sensitive if below-average in node 2 and above-average in node 3.
+We performed a one-tailed, unequal-variance t-test to determine differences in elevation between the two groups [@Welch1947].
+
+Organizing the datasets relied on R packages ClimMobTools [@climmobtools] and tidyverse [@tidyverse].
+Statistical analysis was performed using packages [...].
+Pearson correlations across station and on-farm traits were visualized with corrplot [@corrplot].
+Forward variable selection depended on packages[...].
+Genomic prediction was performed with packages [...].
+For geospatial analysis we used the packages [...].
+
+## Data and code availability
+
+Full data is available through Dataverse [**create repo with Bioversity Dataverse; put the GitHub link in the description**].
+
+\newpage
+
+# Display items
+
+```{r, fig.cap="**Fig.
+1.** Centralized trials (**A**) derives recommendations from breeders evaluation and some participatory assessments which are geographically biased.
+We propose an approach (**B**) where breeders and farmers select the best candidate genotypes to be tested on-farm in small, blinded and randomized sets of genotypes producing scalable solutions that can be linked to environmental data and the interactions between genotypes and environment.", fig.margin=TRUE, echo=FALSE}
 knitr::include_graphics("../output/scheme/Fig1.png")
@@
 
-# Part 2
+\newpage
 
-Farmer evaluation traits can predict metric traits fairly well, depending on their derivation.
-Overall, spike, and GY observed in first year have a similar performance in predicting GY in the second year (Fig.
-2A).
-Farmers scores (both overall and spike) well predict TGW in 2013.
-There is a gender interaction, especially in overall, as arguably men and women have a different interpretation of the overall appreciation of varieties.
-When men and women criteria are contrasting, the prediction power decreases (see overall, overall.f and overall.m over GY in 2013).
-When they are aligned, the prediction power increases (see overall, overall.f and overall.m over DF in 2013; farmers are very worried about late maturation of varieties).
-
-```{r echo=FALSE, fig.cap="Fig. 2. R", fig.height=1, fig.margin=TRUE, fig.width=1}
-knitr::include_graphics("../output/gp/Fig2.png")
+```{r, fig.cap="**Fig.
+2.** Trait values and predictions on-station and in crowdsourcing fields.
+(A) Broad sense heritabilities for PVS overall evaluation compared with those of yield in stations and across stations and years.
+PVS overall is highly heritable, and is influenced by gender and location.
+(B) Spearman correlations among traits.
+Non-significant correlations are not shown; in the upper triangle of the symmetric matrix, significance is corrected for multiple testing.
+In the lower triangle, significance is considered on a single-test basis.
+(C) Bar plot representing accuracy of on-station and across season genomic predictions.
+Predicted traits are reported on the x axis, and predictors are coded by colors according to the legend to the right.
+(D) Bar plot representing accuracy of genomic prediction when using station data as a predictor of farmer fields’ data.
+(E) Bar plot representing accuracy of genomic prediction when using station data as a predictor of farmer fields’ data, limiting the dataset to landraces originally sampled in locations with minimum day temperature (minDT) lower than the splitting value.
+(F) As in previous panel, but limiting the dataset to landraces originally sampled in locations with minimum day temperature (minDT) higher than the splitting value.", echo=FALSE, fig.height=1, fig.margin=TRUE, fig.width=1}
+knitr::include_graphics("../output/gp/Figure2.png")
 ```
 
-# Part 3
+\newpage
 
-```{r, fig.margin=TRUE, echo=FALSE, fig.cap="Fig.
-3.
-Expected yield gain of adopting variety recommendations based on (L) location model, (E) environmental model and (GxE) genotype by environmental model.", fig.height=5, fig.width=5}
+```{r, fig.cap="**Fig.
+3.**.
+Model performance and modelled yield gain for durum wheat genotypes in Ethiopia based on farmers' evaluation in a citizen science method.
+(**A**) Goodness of fit (pseudo-R^2^) of Plackett-Luce trees determined with 100-fold cross-validation for the intercept-only (IO), location (L), environmental (E) and genotype by environment (GxE) models.
+(**B**) Expected yield gain of adopting variety recommendations based on location (L), environmental (E) and genotype by environment (GxE) models across growing seasons in Ethiopia.
+Red line represents the baseline for expected yield of recommended genotypes based on the Wheat Atlas.
+(**C**) Predicted yield gain across 15 growing seasons (2001 to 2015) in Ethiopia for the environmental (E) and genotype by environment (GxE) model.", fig.margin=TRUE, echo=FALSE, fig.height=5, fig.width=5}
 knitr::include_graphics("../output/yield_gain/yield_gain.png")
@@
 
-# Methods
+\newpage
+# Acknowledgements
+We thank all farmers who evaluated the genotypes.
+**[Others, Seeds4Needs??
+Other Sant'Anna project contracts??]**.
+This work was implemented as part of the CGIAR Research Program on Climate Change, Agriculture and Food Security (CCAFS), which is carried out with support from the CGIAR Trust Fund and through bilateral funding agreements.
+For details please visit https://ccafs.cgiar.org/donors.
+The views expressed in this document cannot be taken to reflect the official opinions of these organizations.
+
+# Author contributions
+MDA, JvE, JP, KdS, SØS, JLJ and MEP designed research; CF, YG, DM, BFM, performed field research; MDA, JP and KdS analyzed the data; [...]
+
+# Conflict of interests
+The authors declare no conflict of interests.
 
 # References
-
-# Display items

### 2019-10-18 script/0.get.data.tricot.R

@@ -1,79 +1,0 @@
-options(stringsAsFactors = F)
-
-library(plyr)
-library(RCurl)
-library(lme4)
-library(bestNormalize)
-
-wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/0.get.data"
-setwd(wd)
-
-#get in raw data from GITHUB
-url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/durumwheat.csv")
-csd<-read.csv(text=url)
-csd<-csd[order(csd[,"genotype"]),]
-
-#get a metric about getting first for each variety
-bygen<-split(csd, csd[,"genotype"])
-#set df as output
-out.csd<-matrix(ncol=5, nrow=length(bygen))
-
-for (i in 1:length(bygen)){
-  tmp<-bygen[[i]]
-  
-  #get propotion of winning
-  wins<-length(which(tmp[,"farmer_rank"]==1))
-  odds<-wins/nrow(tmp)
-  
-  #get mean estimates
-  gy<-mean(tmp[,"gy_gm"], na.rm=T)
-  sl<-mean(tmp[,"mean_sl"], na.rm=T)
-  nsps<-mean(tmp[,"mean_seed_no_spike"], na.rm=T)
-  net<-mean(tmp[,"mean_tn"], na.rm=T)
-  
-  #bring everything in the output
-  out.csd[i,]<-c(odds, gy, sl, nsps, net)
-}
-
-colnames(out.csd)<-c("win_prop", "mean_gy", "sl", "nsps", "net")
-rownames(out.csd)<-unique(csd[,"genotype"])
-
-out.csd<-as.data.frame(out.csd)
-
-#get gy from a mixed model accounting for farmer and genotypes
-mod<-lmer(gy_gm  ~  (1|genotype) + (1|farmer) + (1|year), data=csd)
-idblup<-ranef(mod)
-gy<-idblup$genotype
-out.csd<-merge(out.csd, gy, by="row.names", all.x=T)
-rownames(out.csd)<-out.csd[,1]
-out.csd<-out.csd[,-1]
-colnames(out.csd)[ncol(out.csd)]<-"gy_blup"
-
-#now add in data from Kaue's models
-url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/log-abilities/log-abilities.csv")
[…]

### 2019-10-18 script/03_PL_model.R

@@ -58,7 +58,7 @@ am[1:10, 1:10]
 
 am[genitems,] <- gene
 
-am[1:10, 1:10]
+am[1:10, ]
 
 am <- t(am)
 
@@ -122,6 +122,7 @@ gen <- crossvalidation(G ~ minDT_veg + minNT_veg,
                        drop.folds = out,
                        minsize = minsize,
                        alpha = alpha,
+                       gamma = gamma,
                        normal = prior)
 
 print(gen)
@@ -138,6 +139,5 @@ print(loc)
 
 save(null, env, gen, loc,
      file = paste0(output, "models.rda"))
-
 
 

### 2019-10-18 script/04_gain_gy_reliability.R

@@ -2,6 +2,7 @@ library("tidyverse")
 library("magrittr")
 library("janitor")
 library("svglite")
+library("gridExtra")
 library("gosset")
 library("PlackettLuce")
 
@@ -67,99 +68,237 @@ baseline <- mean(baseline[["gy"]])
 #.......................................
 # Gain ####
 # gy by taking the best three varieties from Plackett-Luce models
+# 
+# names(coord) <- c("lon","lat")
+# coord %<>% 
+#   mutate(xy = lon + lat,
+#          yx = lon - lat)
+# 
+# loc_pred <- predict(loc, newdata = coord)
+# loc_gy <- reliability_gy(loc_pred, gy, weights = rep(1/3, 3))
+# 
+# set.seed(1234)
+# loc_gy <- runif(length(loc_gy), baseline-0.15, max(loc_gy))
+# 
+# 
+# env_gy <- NULL
+# for(i in seq_along(climatology)){
+#   print(i)
+#   y <- predict(env, newdata = climatology[[i]])
+#   y <- reliability_gy(y, gy, weights =  c(.7, .15, .15))
+#   y <- data.frame(y = y, sample = paste0("S",i), r = 1:length(y))
+#   env_gy <- rbind(env_gy, y)
+# }
+# 
+# gen_gy <- NULL
+# for(i in seq_along(climatology)){
+#   print(i)
+#   y <- predict(gen, newdata = climatology[[i]])
+#   y <- reliability_gy(y, gy, weights = c(.7, .15, .15))
+#   y <- data.frame(y = y, sample = paste0("S",i), r = 1:length(y))
+#   gen_gy <- rbind(gen_gy, y)
+# }
+# 
+# 
+# save(gen_gy, loc_gy, env_gy, baseline, gy, output,
+#      file = paste0(output, "predictions.rda"))
 
-names(coord) <- c("lon","lat")
-coord %<>% 
-  mutate(xy = lon + lat,
-         yx = lon - lat)
-
-loc_pred <- predict(loc, newdata = coord)
-loc_gy <- reliability_gy(loc_pred, gy, weights = rep(1/3, 3))
-
-set.seed(1234)
-loc_gy <- runif(length(loc_gy), baseline-0.15, max(loc_gy))
-
-
-
-
-env_gy <- NULL
-for(i in seq_along(climatology)){
-  print(i)
-  y <- predict(env, newdata = climatology[[i]])
[…]

### 2019-10-18 script/06_performance_genotype_locations.R

@@ -26,11 +26,11 @@ ind %<>%
 #.....................................
 #.....................................
 # create PlackettLuce rankings ####
-G <- to_rankings(data = df,
-                 items = "genotype",
-                 input = "farmer_rank",
-                 id = "id", 
-                 grouped.rankings = TRUE)
+G <- rank_PL(data = df,
+             items = "genotype",
+             input = "farmer_rank",
+             id = "id",
+             grouped = TRUE)
 
 #.................................................
 #.................................................

### 2019-10-18 script/1.rrblup.R

@@ -1,246 +1,0 @@
-######################
-#this script performs GP with rrBLUP
-#####################
-options(stringsAsFactors = F)
-library(ggplot2)
-library(rrBLUP)
-library(reshape2)
-library(plyr)
-library(RCurl)
-
-#wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/1.rrBLUP"
-#setwd(wd)
-
-########  SNP DATA  ########
-#get in genotypes and clean them
-hmp<-read.delim("data/diversity.panel.svevo.positions.q10.hmp")
-snp.pos<-hmp[,1:11]
-geno<-hmp[,12:ncol(hmp)]
-rownames(geno)<-hmp[,1]
-colnames(geno)<-sub("^X", "", colnames(geno))
-
-#set a function converting genotypic data in the proper format to run rrBLUP
-convert<-function(x){
-  tmp<-as.character(x)
-  tmp[tmp=="N"]<-NA
-  tmp[tmp  %in% c("R","Y","S","W","K","M") ]<-0
-  uniqus<-sort(unique(tmp))
-  uniqus<-uniqus[!uniqus %in% c("0", NA)]
-  if(length(uniqus)>1){
-    tmp[tmp==uniqus[1]]<- 1
-    tmp[tmp==uniqus[2]]<- -1
-  }    
-  tmp[tmp==uniqus[1]]<-1
-  #overwrite
-  return(tmp)
-}
-
-#convert snps in -1,0,1 and NA
-geno2<-apply(geno, 1, convert)
-geno2<-data.frame(geno2)
-rownames(geno2)<-colnames(geno)
-geno2<-apply(geno2, 2, as.numeric)
-
-#perform imputation with rrBLUP
-imputed<-A.mat(geno2, max.missing=0.5, impute.method="mean", return.imputed=T)
-failrate<-apply(imputed$imputed,2, function(x) length(which(is.na(x))))
-todrop<-which(failrate>0)
-if(length(todrop)>1){
-  imputedcl<-imputed$imputed[,-which(failrate>0)]
-} else {
-  imputedcl<-imputed$imputed
-}
-rownames(imputedcl)<-colnames(geno)
-
-########  PHENOTYPIC DATA  ########
-#get in BLUP values and clean them
-phenos<-read.csv("data/Et_diversity_panel_ALL-TRAITS.BLUPS.csv")
[…]

### 2019-10-18 script/GP_0.get.data.tricot.GP.R

@@ -1,0 +1,430 @@
+# ..........................................
+# ..........................................
+# Combine genomic data with farmer metrics for genomic prediction
+# Matteo Dell'Acqua
+# ..........................................
+# ..........................................
+
+options(stringsAsFactors = F)
+
+library(plyr)
+library(RCurl)
+library(lme4)
+library(bestNormalize)
+library(rrBLUP)
+
+##################
+#DIVERSITY PANEL DATA
+##################
+
+#load in the diversity panel datasets
+load("data/diversity.panel.data.gp.Rdata")
+
+#################
+#SNP data
+
+#set a function converting genotypic data in the proper format to run rrBLUP
+convert<-function(x){
+  tmp<-as.character(x)
+  tmp[tmp=="N"]<-NA
+  tmp[tmp  %in% c("R","Y","S","W","K","M") ]<-0
+  uniqus<-sort(unique(tmp))
+  uniqus<-uniqus[!uniqus %in% c("0", NA)]
+  if(length(uniqus)>1){
+    tmp[tmp==uniqus[1]]<- 1
+    tmp[tmp==uniqus[2]]<- -1
+  }    
+  tmp[tmp==uniqus[1]]<-1
+  #overwrite
+  return(tmp)
+}
+
+#convert snps in -1,0,1 and NA
+geno2<-apply(geno, 1, convert)
+geno2<-data.frame(geno2)
+geno2<-apply(geno2, 2, as.numeric)
+rownames(geno2)<-colnames(geno)
+
+#perform imputation with rrBLUP
+imputed<-A.mat(geno2, max.missing=0.5, impute.method="mean", return.imputed=T)
+failrate<-apply(imputed$imputed,2, function(x) length(which(is.na(x))))
+todrop<-which(failrate>0)
+if(length(todrop)>1){
+  imputedcl<-imputed$imputed[,-which(failrate>0)]
+} else {
+  imputedcl<-imputed$imputed
+}
+rownames(imputedcl)<-colnames(geno)
+
+output <- "processing/GP/"
+dir.create(output, showWarnings = FALSE, recursive = TRUE)
+
+save(imputedcl, file= paste0(output, "genoytpic.data.rrBLUP.Rdata"))
+
+######################
+#calculate BLUPs for metric traits
+
[…]

### 2019-10-18 script/GP_1.rrblup.GP.R

@@ -1,0 +1,274 @@
+# ..........................................
+# ..........................................
+# Genomic prediction with rrBLUP
+# Matteo Dell'Acqua
+# ..........................................
+# ..........................................
+
+options(stringsAsFactors = F)
+library(ggplot2)
+library(rrBLUP)
+library(reshape2)
+library(plyr)
+library(RCurl)
+library(corrplot)
+library(psych)
+
+#wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/1.rrBLUP"
+#setwd(wd)
+
+####load SNP data
+load("processing/GP/genoytpic.data.rrBLUP.Rdata")
+
+####load diversity panel phentoypes
+#load("../0.get.data/blup.diversity.panel.Rdata")
+mts<-read.delim("../../data/jesse_elaboration/diversity.panel.BLUPs.met.csv", sep=",")
+fms<-read.delim("../../data/jesse_elaboration/diversity.panel.BLUPs.farm.csv", sep=",")
+colnames(fms)<-sub("TILLER", "tillering", colnames(fms))
+colnames(fms)<-sub("OVERALL", "overall", colnames(fms))
+colnames(fms)<-sub("SPIKE", "spike", colnames(fms))
+colnames(fms)<-sub("EARLINESS", "earliness", colnames(fms))
+colnames(fms)<-sub(".F$", ".W$", colnames(fms))
+phenos<-merge(mts, fms, by="ID")
+phenos[,1]<-sub("^ID_", "", phenos[,1])
+rownames(phenos)<-phenos[,1]
+phenos<-phenos[,-1]
+
+####load crowdsourcing data
+load("../0.get.data/cs.data.estimates.Rdata")
+
+####load hot tolerance info
+url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/heat_tolerance/genotype_groups.csv")
+hottol<-read.csv(text=url)
+
+#start by bringing together crowdsourcing data and phenotypes
+pheno<-merge(phenos, out.csd, by="row.names", all.x=T)
+rownames(pheno)<-pheno[,1]
+pheno<-pheno[,-1]
+
+#make sure no bread wheat samples are included
+todrop<-grep("_B", rownames(pheno))
[…]

### 2019-10-18 script/GP_3.extract.infos.GP.R

@@ -1,0 +1,299 @@
+######################
+#this script looks into predictions and extracts informations
+#####################
+
+options(stringsAsFactors = F)
+wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/2.extract.info"
+setwd(wd)
+
+library(ggplot2)
+library(plyr)
+library(reshape2)
+library(ggpubr)
+library(psych)
+library(ggcorrplot)
+
+#start getting in heritability values
+h2<-read.delim("../../data/jesse_elaboration/diversity.panel.h2.met.csv", sep=",")
+h2$code<-paste(h2[,"trait"], h2[,"year"], h2[,"location"], sep="_")
+
+#list output files in rrBLUP folder
+filen<-list.files("../1.rrBLUP/")
+filen<-filen[grep("GP.Rdata", filen)]
+
+#get only those with Jesse's BLUPs
+filen<-filen[grep("jesse", filen)]
+
+#load the results of GP according to listed files
+rslist<-list()
+for(i in 1:length(filen)){
+  load(paste0("../1.rrBLUP/",filen[i]))
+  rslist[[i]]<-predresults
+}#for i
+names(rslist)<-filen
+names(rslist)
+
+#make an output list
+outlist<-list()
+
+#create output vector holding ALL trait nanmes
+alltraitnames<-c()
+
+#summarize results by trait by prediction
+for(pl in 1:length(rslist)){
+  
+  tmppred<-rslist[[pl]]
+  
+  gptraits<-colnames(tmppred[[1]])
+  
+  alltraitnames<-c(alltraitnames, gptraits)
+  
+  tbyt<-list()
+  
+  for (i in 1:length(gptraits)){
+    tmp<-lapply(tmppred, function(x) data.frame(x[,i]))
+    tmpout<-do.call(cbind, tmp)
+    rownames(tmpout)<-rownames(tmppred[[1]])
+    colnames(tmpout)<-1:ncol(tmpout)
+    tbyt[[i]]<-tmpout
+    names(tbyt)[i]<-gptraits[i]
+  }
+  
+  outlist[[pl]]<-tbyt
+  
+}#for pl
+names(outlist)<-names(rslist)
+
+#get unique trait names
+alltraitnames<-sort(unique(alltraitnames))
+
+#set up colors matching trait names
+coldf<-data.frame(alltraitnames, col=NA)
[…]

### 2019-10-18 script/GP_3.hot.cold.R

@@ -1,0 +1,65 @@
+######################
+#this script conducts analyses considering adaptation to DT_min_veg
+#####################
+
+options(stringsAsFactors = F)
+wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/3.hot.cold"
+setwd(wd)
+
+library(ggplot2)
+library(plyr)
+library(RCurl)
+library(reshape2)
+
+#load diversity panel phentoypes
+load("../0.get.data/blup.diversity.panel.Rdata")
+#load crowdsourcing data
+load("../0.get.data/cs.data.estimates.Rdata")
+#load hot tolerance info
+url<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/output/heat_tolerance/genotype_groups.csv")
+hottol<-read.csv(text=url)
+#set hot samples and cold samples
+hot.samples<-hottol[which(hottol$group==1),1]
+cold.samples<-hottol[which(hottol$group==2),1]
+
+#add cold/hot info to cs data and dp data
+allph<-merge(phenos, out.csd, by="row.names", all.y=T)
+rownames(allph)<-allph[,1]
+allph[,1]<-"warm"
+allph[which(row.names(allph) %in% cold.samples),1]<-"cold"
+colnames(allph)[1]<-"adapt"
+
+allph$adapt<-as.factor(allph$adapt)
+
+#sort out colds and hots
+colds<-allph[which(allph$adapt=="cold"),2:ncol(allph)]
+warms<-allph[which(allph$adapt=="warm"),2:ncol(allph)]
+
+
+#check some correlations
+cr<-cor(colds, use="complete.obs")
+corrplot(cr)
+
+#make some plots
+boxplot(DF ~ adapt, data=allph)
+t.test(DF ~ adapt, data=allph)
+
+boxplot(GY ~ adapt, data=allph)
+t.test(GY ~ adapt, data=allph)
+
+boxplot(TGW ~ adapt, data=allph)
+t.test(TGW ~ adapt, data=allph)
+
+boxplot(overall ~ adapt, data=allph)
+t.test(TGW ~ adapt, data=allph)
+
+plot(allph$GY, allph$DF)
+
+boxplot(GY_BLUP.cs ~ adapt, data=allph)
+t.test(GY_BLUP.cs ~ adapt, data=allph)
+
+#make violin plots
[…]

### 2019-10-18 script/GP_4.env.space.R

@@ -1,0 +1,87 @@
+######################
+#this script checks the environmental space covered by tested locations
+##################
+
+options(stringsAsFactors = F)
+wd<-"C:/Users/admin/OneDrive - Scuola Superiore Sant'Anna/projects/tricot.genomics/analyses/4.env.space"
+setwd(wd)
+
+library(ggplot2)
+library(plyr)
+library(RCurl)
+library(reshape2)
+library(plotly)
+library(plot3D)
+
+#get station climate
+url1<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/environmental_indices_station.csv")
+stat<-read.csv(text=url1)
+#get summary statistics
+stat<-aggregate(. ~ location, data = stat, FUN = mean)
+rownames(stat)<-stat[,1]
+stat<-stat[,-1]
+stat<-stat[,-c(58:ncol(stat))]
+
+#get cs climate
+url2<-getURL("https://raw.githubusercontent.com/agrobioinfoservices/tricot-genomic/master/data/environmental_indices.csv")
+farm<-read.csv(text=url2)
+rownames(farm)<-farm$id
+farm<-farm[,-c(49:ncol(farm))]
+
+#merge the two dataframes only considering common columns
+stat<-stat[,which(colnames(stat) %in% colnames(farm))]
+farm<-farm[,which(colnames(farm) %in% colnames(stat))]
+stopifnot(all(colnames(stat)==colnames(farm)))
+
+#assemble the dataset and move to PCA
+alldat<-data.frame(rbind(farm, stat))
+colz<-rep("gray", nrow(alldat))
+colz[(length(colz)-1):length(colz)]<-"red"
+
+pca<-prcomp(alldat, scale = T)
+pov<-pca$sdev^2/sum(pca$sdev^2)
+
+plot(pca$x[,1:2], col=colz)
+
+pctoplot<-data.frame(cbind(colz, pca$x[,1:3],alldat))
+pctoplot$pch<- "circle"
+pctoplot$pch[(length(pctoplot$pch)-1):length(pctoplot$pch)]<-"diamond-open"
+pctoplot$cex<-"6"
+pctoplot$cex[(length(pctoplot$cex)-1):length(pctoplot$cex)]<-"8"
+
+#make a 3d pc plot with plotly
+p <- plot_ly(pctoplot, x = ~PC1, y = ~PC2, z = ~PC3,
+             marker = list(color = ~minDT_veg,
+                           symbol = ~pch, 
[…]

### 2019-10-18 script/helper_00_functions.R

@@ -1,5 +1,7 @@
 # compute gain in grain yield based on model predictions
-reliability_gy <- function(object, gy, weights = c(0.5, 0.3, 0.2), sort = "yield") {
+reliability_gy <- function(object, gy, 
+                           weights = c(0.5, 0.3, 0.2), 
+                           sort = "yield") {
   
   if(sum(weights) != 1) {
     stop("Sum of weights is not 1")

### 2019-10-18 script/helper_08_plot_gp_outputs.R

@@ -1,0 +1,55 @@
+library("tidyverse")
+library("magrittr")
+library("reshape2")
+library("svglite")
+library("gridExtra")
+
+# read data produced in script 1.rrblup.R
+load("data/GP.CSdata.Rdata")
+
+#['#b2182b','#ef8a62','#fddbc7','#d1e5f0','#67a9cf','#2166ac']
+#['#762a83','#af8dc3','#e7d4e8','#d9f0d3','#7fbf7b','#1b7837']
+
+# produce some plotting ####
+gptraits <- colnames(predresults[[1]])
+tbyt <- list()
+
+for (i in 1:length(gptraits)) {
+  tmp <- lapply(predresults, function(x)
+    data.frame(x[, i]))
+  tmpout <- do.call(cbind, tmp)
+  rownames(tmpout) <- rownames(predresults[[1]])
+  colnames(tmpout) <- 1:ncol(tmpout)
+  tbyt[[i]] <- tmpout
+  names(tbyt)[i] <- gptraits[i]
+}
+
+nsubs <- ncol(tbyt[[1]])
+
+p <- list()
+
+# reshape the dataframes and plot them
+for (i in 1:length(tbyt)){
+  gptrait <- names(tbyt)[i]
+  tmp <- tbyt[[i]]
+  toplot <- melt(t(tmp))
+  colnames(toplot) <- c("Rep", "Trait", "Accuracy")
+  
+  # get summary stats
+  toplot %<>% 
+    group_by(Trait) %>% 
+    summarise(value = mean(Accuracy, na.rm = TRUE),
+              se = sd(Accuracy) / sqrt(length(Accuracy)))
+  
+  p[[i]]<- ggplot(toplot) +
+    geom_bar(aes(x=Trait, y=value), stat="identity", fill="skyblue", alpha=0.7) +
+    geom_errorbar(aes(x=Trait, ymin=value-se, ymax=value+se), 
+                  width=0.4, colour="black", alpha=0.9, size=0.5) +
+    labs(title=gptrait)+
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+ 
+  
+}
+
+
+p[[5]]
